### 2025-02-25 paper.qmd
__init__

---
title: "The Tail Risk Premium - Extension of Lottery Preference Premium"
format: html

---
## Abstract
The investment strategy centered around the Idiosyncratic Volatility (IVOL) Risk Premium exploits the distinct performance patterns of stocks exhibiting varying levels of firm-specific risk. Unlike systemic risk, which permeates the entire market, idiosyncratic volatility is unique to individual firms and can be effectively diversified within a well-constructed portfolio. Empirical research consistently demonstrates that stocks with low idiosyncratic volatility tend to outperform those with high idiosyncratic volatility, suggesting a negative risk premium for high-IVOL stocks and a positive premium for low-IVOL stocks. This phenomenon has been validated across diverse international markets and controlled for market, size, and book-to-market factors, forming the foundation of this investment strategy.

## Investment Strategy
The core of the strategy involves taking long positions in stocks characterized by low idiosyncratic volatility, which consistently exhibit a positive return premium. These stocks are frequently underpriced relative to their risk profiles, presenting an opportunity for reliable excess returns. Conversely, stocks with high idiosyncratic volatility are systematically mispriced due to behavioral biases such as investor preference for lottery-like payoffs and market frictions that limit arbitrage. By strategically shorting these high-IVOL stocks, investors can exploit their documented pattern of underperformance. This dual approach leverages market inefficiencies to achieve superior risk-adjusted returns.

## Persistence of the Idiosyncratic Risk Premium:
The persistence of the idiosyncratic risk premium is attributed to several factors, including investor behavior, market frictions, and liquidity constraints. High-IVOL stocks attract risk-tolerant investors seeking high returns, but deter conservative investors, creating a premium for stocks with lottery-like features. Market frictions, such as high transaction costs and information asymmetry, further perpetuate this premium by impeding effective arbitrage. Additionally, liquidity constraints necessitate higher compensation for holding assets with high idiosyncratic risk. By comprehensively understanding and exploiting these dynamics, the investment strategy aims to deliver consistent and predictable returns.

## Strategy Details
To effectively harness the Idiosyncratic Volatility (IVOL) Risk Premium, the strategy employs a systematic approach in selecting and managing investments:

### Data Selection:
+ The initial step involves gathering comprehensive data from the WRDS database by Wharton, spanning from 2002 to 2022. This extensive data set includes stock prices, historical returns, and idiosyncratic volatility measures.
+ The portfolio will be rebalanced annually from 2007 to 2022, ensuring a lookback window of five additional years to refine the selection criteria.

### Portfolio Construction
+ Selection Criteria: The strategy identifies stocks based on idiosyncratic volatility and skewness. Stocks exhibiting high idiosyncratic volatility and negative skewness are shortlisted for short positions, while those with low idiosyncratic volatility and positive skewness are chosen for long positions.

+ Long Positions: The strategy takes long positions in stocks with low idiosyncratic volatility, aiming to capitalize on their consistent positive return premium. These stocks are often underpriced relative to their risk profiles, offering reliable opportunities for excess returns.

+ Short Positions: Simultaneously, short positions are established in stocks with high idiosyncratic volatility. These stocks are systematically mispriced due to behavioral biases and market frictions, and they tend to underperform over time. The strategy seeks to profit from their predictable pattern of underperformance.

### Risk Management:
+ Diversification: The portfolio is diversified across multiple stocks to mitigate firm-specific risks. By constructing a well-balanced portfolio, the strategy ensures that the impact of any single stock's volatility is minimized.
+ Rebalancing: Annual rebalancing is conducted to maintain the desired risk exposure and adjust for changes in idiosyncratic volatility and skewness. This regular review allows the strategy to adapt to evolving market conditions and optimize returns.

### Performance Monitoring:
+ The strategy's performance is continuously monitored against relevant benchmarks to ensure alignment with its objectives. Key performance indicators, such as risk-adjusted returns and volatility measures, are tracked to assess the strategy's effectiveness.
+ Adjustments: Based on performance analysis, adjustments may be made to the portfolio composition, selection criteria, or rebalancing frequency to enhance the strategy's overall performance and resilience to market fluctuations.

### 2025-02-25 functions.py

@@ -1,0 +1,9 @@
+import pandas as pd
+import numpy as np
+
+def rebalance():
+    x=1
+
+def skew():
+    x=1
+

### 2025-02-25 portfolio.py

@@ -1,0 +1,5 @@
+def portfolio():
+    x=1
+
+def visualize():
+    x=1

### 2025-02-25 portfolio.py
Added V1 Screens + V2 Init

@@ -1,5 +1,94 @@
-def portfolio():
-    x=1
+import pandas as pd
+import numpy as np
+from scipy.stats import skew, kurtosis
 
-def visualize():
-    x=1
+from screens import v1, v2, rebalance
+
+def data():
+    dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
+    df = pd.read_csv(dataFile)
+
+    df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
+    df = df.sort_values(by='DATE', ascending=True)
+    df['MONTH'] = df['DATE'].dt.month
+    df['YEAR'] = df['DATE'].dt.year
+
+    def convert_to_numeric(column):
+        return pd.to_numeric(column.str.replace('%', ''), errors='coerce')
+    
+    df['RET'] = convert_to_numeric(df['RET']) / 100
+    df['IVOL'] = convert_to_numeric(df['ivol']) / 100
+
+    df = df[['DATE', 'MONTH', 'YEAR', 'TICKER', 'RET', 'IVOL']]
+
+    return df
+
+def metaData(select=None):
+    df = data()
+    if select == 'tickerCount':
+        ticker_count = df['TICKER'].nunique()
+        return ticker_count
+    
+    elif select == 'tickers':
+        tickers = list(df['TICKER'].unique())
+        return tickers
+    
+    elif select == 'years':
+        years = list(df['YEAR'].unique())
+        return years
+
+def portfolio():
+    df = data()
+    v1_list = [] ## Cols: Year, LongPosV1, ShortPosV1, LongRetV1, ShortRetV1, LongSkewV1, ShortSkewV1
+    
+    rebalance_period = metaData(select='years')
+
+    ## V1->V2 Screen (Extreme IVOL High-Low) -> (Extreme SKEW Positive-Negative)
+    for period in rebalance_period:
+        df_back = df[(df['YEAR'] <= (period - 1)) & (df['YEAR'] >= (period - 6))].copy()  ## Five Year Lookback Window
+        df_back['PERIOD'] = period ## Rebalance Year
+        
+        stats_df = df_back.groupby('TICKER')['IVOL'].agg(
+            AVE='mean',
+            SKEW='skew'
+            ).reset_index()
+        
+        return_df = df[df['YEAR'] == period]
+        return_df = df_back.groupby('TICKER')['RET'].agg(
[…]

### 2025-02-25 screens.py

@@ -1,9 +1,12 @@
 import pandas as pd
 import numpy as np
 
-def rebalance():
+def v1():
     x=1
 
-def skew():
+def v2():
+    x=1
+
+def rebalance():
     x=1
 

### 2025-02-26 paper.qmd
Added Paper
Updates To Paper

@@
 format: html
+author: "Brayden Boyko, Roan Henderson, Andrew Pettifor"
+jupyter: 
+    kernel: python3
 
@@
 + Adjustments: Based on performance analysis, adjustments may be made to the portfolio composition, selection criteria, or rebalancing frequency to enhance the strategy's overall performance and resilience to market fluctuations.
+
+```{Python}
+#| echo: false
+#| warning: false
+
+```
+
+## Performance Evaluations
+A variety of benchmark statistics will be included in our portfolio testing:
++ Annual / Monthly Return Values/Charts For 4 Portfolios
++ Describe Stock Breakdown For Four Portfolios (Skew, IVOL)
++ V1 & V2 Screening

### 2025-02-26 portfolio.py

@@ -1,12 +1,12 @@
 import pandas as pd
 import numpy as np
-from scipy.stats import skew, kurtosis
 
 from screens import v1, v2, rebalance
 
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile)
+    dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
+    df = pd.read_csv(dataFile2)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)
@@ -43,6 +43,8 @@ def portfolio():
     
     rebalance_period = metaData(select='years')
 
+    ### + Add Random Sample
+
     ## V1->V2 Screen (Extreme IVOL High-Low) -> (Extreme SKEW Positive-Negative)
     for period in rebalance_period:
         df_back = df[(df['YEAR'] <= (period - 1)) & (df['YEAR'] >= (period - 6))].copy()  ## Five Year Lookback Window
@@ -50,8 +52,11 @@ def portfolio():
         
         stats_df = df_back.groupby('TICKER')['IVOL'].agg(
             AVE='mean',
-            SKEW='skew'
             ).reset_index()
+        
+        stats2_df = df_back.groupby('TICKER')['RET'].agg(
+            SKEW='skew'
+        ).reset_index()
         
         return_df = df[df['YEAR'] == period]
         return_df = df_back.groupby('TICKER')['RET'].agg(
@@ -59,6 +64,7 @@ def portfolio():
             ).reset_index()
         
         df_back = df_back.merge(stats_df, on='TICKER', how='left')
+        df_back = df_back.merge(stats2_df, on='TICKER', how='left')
         df_back = df_back.merge(return_df, on='TICKER', how='left')
         df_back = df_back.drop_duplicates(subset='TICKER')
         df_back = df_back.dropna()
@@ -83,8 +89,7 @@ def portfolio():
     v1_df = v1_df[v1_df['Year'] >= 2007] ## Rebalances Begin In 2007
     
     print(v1_df)
-
-
+    v1_df.to_csv(r"c:\Users\brayd\Downloads\test.csv")
 
 
 

### 2025-02-26 portfolio.py
Added V2 Filter

@@ -6,7 +6,7 @@ from screens import v1, v2, rebalance
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile2)
+    df = pd.read_csv(dataFile)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)
@@ -87,9 +87,29 @@ def portfolio():
     v1_df = v1_df.pivot(index=['Year', 'TICKER'], columns='Position', values=['T_RETURN', 'SKEW']).reset_index()
     v1_df.columns = ['Year', 'TICKER', 'LongRetV1', 'ShortRetV1', 'LongSkewV1', 'ShortSkewV1']
     v1_df = v1_df[v1_df['Year'] >= 2007] ## Rebalances Begin In 2007
+
+    rebalance_years = list(v1_df['Year'].unique())
+
+    v2_list = []
+    for year in rebalance_years:
+        v2_df = v1_df[v1_df['Year'] == year]
+        
+        v2_long = v2_df[['Year', 'TICKER','LongRetV1', 'LongSkewV1']].dropna()
+        v2_long.columns = ['Year', 'TICKER', 'LongRetV2', 'LongSkewV2']
+        v2_long = v2_long.sort_values(by='LongSkewV2', ascending=True)
+        v2_long = v2_long.tail(30)
+
+        v2_short = v2_df[['Year', 'TICKER','ShortRetV1', 'ShortSkewV1']].dropna()
+        v2_short.columns = ['Year', 'TICKER', 'ShortRetV2', 'ShortSkewV2']
+        v2_short = v2_short.sort_values(by='ShortSkewV2', ascending=True)
+        v2_short = v2_short.head(30)
+
+        combined = pd.concat([v2_long, v2_short], axis=0)
+        v2_list.append(combined)
+
+    v2_df = pd.concat(v2_list, ignore_index=True)
     
-    print(v1_df)
-    v1_df.to_csv(r"c:\Users\brayd\Downloads\test.csv")
+    print(v2_df)
 
 
 

### 2025-02-26 portfolio.py
Added Annual portfolio returns

@@ -111,7 +111,28 @@ def portfolio():
     
     print(v2_df)
 
+    ## Long-Short Base Startegy (L_Base/S_Base) ---> Long-Short Portfolio Startegy (L_Port/S_Port)
 
+    L_Base = v1_df[['Year', 'TICKER', 'LongRetV1', 'LongSkewV1']].dropna()
+    L_Base = L_Base.groupby('Year')['LongRetV1'].agg(LongV1Ret='mean').reset_index()
+
+    S_Base = v1_df[['Year', 'TICKER', 'ShortRetV1', 'ShortSkewV1']].dropna()
+    S_Base = S_Base.groupby('Year')['ShortRetV1'].agg(ShortV1Ret='mean').reset_index()
+    S_Base['ShortV1Ret'] = S_Base['ShortV1Ret'] * -1 ## Make Returns Inverse
+
+    L_Port = v2_df[['Year', 'TICKER', 'LongRetV2', 'LongSkewV2']].dropna()
+    L_Port = L_Port.groupby('Year')['LongRetV2'].agg(LongV2Ret='mean').reset_index()
+
+    S_Port = v2_df[['Year', 'TICKER', 'ShortRetV2', 'ShortSkewV2']].dropna()
+    S_Port = S_Port.groupby('Year')['ShortRetV2'].agg(ShortV2Ret='mean').reset_index()
+    S_Port['ShortV2Ret'] = S_Port['ShortV2Ret'] * -1 ## Make Returns Inverse
+
+    port = L_Base.merge(L_Port, how='left', on='Year')
+    port = port.merge(S_Base, how='left', on='Year')
+    port = port.merge(S_Port, how='left', on='Year')
+    
+
+    print(port)
 
 
 portfolio()

### 2025-02-26 portfolio.py
Updates To Short Position Calculation

@@ -108,8 +108,6 @@ def portfolio():
         v2_list.append(combined)
 
     v2_df = pd.concat(v2_list, ignore_index=True)
-    
-    print(v2_df)
 
     ## Long-Short Base Startegy (L_Base/S_Base) ---> Long-Short Portfolio Startegy (L_Port/S_Port)
 
@@ -118,14 +116,14 @@ def portfolio():
 
     S_Base = v1_df[['Year', 'TICKER', 'ShortRetV1', 'ShortSkewV1']].dropna()
     S_Base = S_Base.groupby('Year')['ShortRetV1'].agg(ShortV1Ret='mean').reset_index()
-    S_Base['ShortV1Ret'] = S_Base['ShortV1Ret'] * -1 ## Make Returns Inverse
+    S_Base['ShortV1Ret'] = (S_Base['ShortV1Ret'] * -1).clip(lower=-1) ## Make Returns Inverse
 
     L_Port = v2_df[['Year', 'TICKER', 'LongRetV2', 'LongSkewV2']].dropna()
     L_Port = L_Port.groupby('Year')['LongRetV2'].agg(LongV2Ret='mean').reset_index()
 
     S_Port = v2_df[['Year', 'TICKER', 'ShortRetV2', 'ShortSkewV2']].dropna()
     S_Port = S_Port.groupby('Year')['ShortRetV2'].agg(ShortV2Ret='mean').reset_index()
-    S_Port['ShortV2Ret'] = S_Port['ShortV2Ret'] * -1 ## Make Returns Inverse
+    S_Port['ShortV2Ret'] = (S_Port['ShortV2Ret'] * -1).clip(lower=-1) ## Make Returns Inverse
 
     port = L_Base.merge(L_Port, how='left', on='Year')
     port = port.merge(S_Base, how='left', on='Year')
@@ -133,6 +131,7 @@ def portfolio():
     
 
     print(port)
+    return port
 
 
 portfolio()

### 2025-02-26 portfolio.py
Added Position Tickers To Annual Portfolio Output

@@ -6,7 +6,7 @@ from screens import v1, v2, rebalance
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile)
+    df = pd.read_csv(dataFile2)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)
@@ -113,22 +113,32 @@ def portfolio():
 
     L_Base = v1_df[['Year', 'TICKER', 'LongRetV1', 'LongSkewV1']].dropna()
     L_Base = L_Base.groupby('Year')['LongRetV1'].agg(LongV1Ret='mean').reset_index()
+    tickers = v1_df.groupby('Year')['TICKER'].unique().reset_index(name='LongPosV1')
+    L_Base = L_Base.merge(tickers, on='Year', how='left')
 
     S_Base = v1_df[['Year', 'TICKER', 'ShortRetV1', 'ShortSkewV1']].dropna()
     S_Base = S_Base.groupby('Year')['ShortRetV1'].agg(ShortV1Ret='mean').reset_index()
     S_Base['ShortV1Ret'] = (S_Base['ShortV1Ret'] * -1).clip(lower=-1) ## Make Returns Inverse
+    tickers = v1_df.groupby('Year')['TICKER'].unique().reset_index(name='ShortPosV1')
+    S_Base = S_Base.merge(tickers, on='Year', how='left')
 
     L_Port = v2_df[['Year', 'TICKER', 'LongRetV2', 'LongSkewV2']].dropna()
     L_Port = L_Port.groupby('Year')['LongRetV2'].agg(LongV2Ret='mean').reset_index()
+    tickers = v2_df.groupby('Year')['TICKER'].unique().reset_index(name='LongPosV2')
+    L_Port = L_Port.merge(tickers, on='Year', how='left')
 
     S_Port = v2_df[['Year', 'TICKER', 'ShortRetV2', 'ShortSkewV2']].dropna()
     S_Port = S_Port.groupby('Year')['ShortRetV2'].agg(ShortV2Ret='mean').reset_index()
     S_Port['ShortV2Ret'] = (S_Port['ShortV2Ret'] * -1).clip(lower=-1) ## Make Returns Inverse
[…]

### 2025-02-27 monthly_returns.py
Added Monthly Returns

@@ -1,0 +1,24 @@
+from portfolio import portfolio, data
+
+def monthly_returns(Pos=None, Port=None):
+    """
+    ## Generate a DataFrame With Monthly Returns For Specified Portfolio 
+    ---
+    + Pos: Position Argument ("Long", "Short")
+    + Port: Portfolio ("V1", "V2")
+    """
+
+    df = data()
+    port = portfolio()
+    select = f"{Pos}Pos{Port}"
+
+    print(select)
+    
+    port = port[[]]
+
+
+
+    print(portfolio)
+
+
+monthly_returns(Pos="Long", Port="V1")

### 2025-02-27 portfolio.py

@@ -1,8 +1,6 @@
 import pandas as pd
 import numpy as np
 
-from screens import v1, v2, rebalance
-
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
@@ -137,13 +135,5 @@ def portfolio():
     port = port.merge(S_Base, how='left', on='Year')
     port = port.merge(S_Port, how='left', on='Year')
 
-
-
-
     print(port)
     return port
-
-
-portfolio()
-
-

### 2025-02-27 screens.py

@@ -1,12 +1,0 @@
-import pandas as pd
-import numpy as np
-
-def v1():
-    x=1
-
-def v2():
-    x=1
-
-def rebalance():
-    x=1
-

### 2025-02-27 monthly_returns.py
Added Monthly Returns For All Portfolios

@@ -1,4 +1,5 @@
 from portfolio import portfolio, data
+import pandas as pd
 
 def monthly_returns(Pos=None, Port=None):
     """
@@ -11,14 +12,29 @@ def monthly_returns(Pos=None, Port=None):
     df = data()
     port = portfolio()
     select = f"{Pos}Pos{Port}"
-
-    print(select)
     
-    port = port[[]]
+    port = port[['Year', f'{select}']]
 
+    df = df.merge(port, how='left', left_on='YEAR', right_on='Year').dropna()
+    df = df.explode(f'{select}')
+    df = df[df['TICKER'] == df[f'{select}']]
+    df['PORTFOLIO'] = select
 
+    df = df[['DATE', 'YEAR', 'TICKER', 'RET', 'IVOL', 'PORTFOLIO']]
+
+    return df
+
+def portfolio_returns():
+
+    L_v1 = monthly_returns(Pos="Long", Port="V1")
+    S_v1 = monthly_returns(Pos="Short", Port="V1")
+    L_v2 = monthly_returns(Pos="Long", Port="V1")
+    S_v2 = monthly_returns(Pos="Short", Port="V2")
+
+    portfolio = pd.concat([L_v1, S_v1, L_v2, S_v2])
+    portfolio = portfolio.groupby(['DATE', 'PORTFOLIO']).agg({'RET': 'mean', 'IVOL': 'mean'}).reset_index()
 
     print(portfolio)
+    return portfolio
 
-
-monthly_returns(Pos="Long", Port="V1")
+portfolio_returns()

### 2025-02-27 portfolio.py

@@ -4,7 +4,7 @@ import numpy as np
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile2)
+    df = pd.read_csv(dataFile)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)
@@ -135,5 +135,4 @@ def portfolio():
     port = port.merge(S_Base, how='left', on='Year')
     port = port.merge(S_Port, how='left', on='Year')
 
-    print(port)
     return port

### 2025-02-27 monthly_returns.py
Added Basic Charts

@@ -28,13 +28,12 @@ def portfolio_returns():
 
     L_v1 = monthly_returns(Pos="Long", Port="V1")
     S_v1 = monthly_returns(Pos="Short", Port="V1")
-    L_v2 = monthly_returns(Pos="Long", Port="V1")
+    L_v2 = monthly_returns(Pos="Long", Port="V2")
     S_v2 = monthly_returns(Pos="Short", Port="V2")
 
     portfolio = pd.concat([L_v1, S_v1, L_v2, S_v2])
     portfolio = portfolio.groupby(['DATE', 'PORTFOLIO']).agg({'RET': 'mean', 'IVOL': 'mean'}).reset_index()
 
-    print(portfolio)
     return portfolio
 
 portfolio_returns()

### 2025-02-27 visuals.py

@@ -1,0 +1,37 @@
+from monthly_returns import portfolio_returns
+
+import pandas as pd
+import matplotlib.pyplot as plt
+
+def monthly_ret():
+    portfolio = portfolio_returns()
+
+    plt.figure(figsize=(12, 6))
+    for port in portfolio['PORTFOLIO'].unique():
+        data = portfolio[portfolio['PORTFOLIO'] == port]
+        plt.plot(data['DATE'], data['RET'], label=port)
+
+    plt.xlabel("Date")
+    plt.ylabel("Return")
+    plt.title("Monthly Portfolio Returns Over Time")
+    plt.legend()
+    plt.grid(True)
+    plt.show()
+
+
+def monthly_ivol():
+    portfolio = portfolio_returns()
+
+    plt.figure(figsize=(12, 6))
+    for port in portfolio['PORTFOLIO'].unique():
+        data = portfolio[portfolio['PORTFOLIO'] == port]
+        plt.plot(data['DATE'], data['IVOL'], label=port)
+
+    plt.xlabel("Date")
+    plt.ylabel("IVOL")
+    plt.title("Monthly Portfolio Idiosyncratic Volatility Over Time")
+    plt.legend()
+    plt.grid(True)
+    plt.show()
+
+monthly_ivol()

### 2025-02-27 portfolio.py
Added Cumulative Returns

@@ -4,7 +4,7 @@ import numpy as np
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile)
+    df = pd.read_csv(dataFile2)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)

### 2025-02-27 visuals.py

@@ -1,6 +1,7 @@
 from monthly_returns import portfolio_returns
 
 import pandas as pd
+import numpy as np
 import matplotlib.pyplot as plt
 
 def monthly_ret():
@@ -18,7 +19,6 @@ def monthly_ret():
     plt.grid(True)
     plt.show()
 
-
 def monthly_ivol():
     portfolio = portfolio_returns()
 
@@ -34,4 +34,33 @@ def monthly_ivol():
     plt.grid(True)
     plt.show()
 
-monthly_ivol()
+def monthly_total_return():
+
+    portfolio = portfolio_returns()
+    portfolio['CUM_RET'] = (1 + portfolio['RET']).cumprod()
+
+    plt.figure(figsize=(12, 6))
+    for port in portfolio['PORTFOLIO'].unique():
+        data = portfolio[portfolio['PORTFOLIO'] == port]
+        plt.plot(data['DATE'], data['CUM_RET'], label=port)
+
+    plt.xlabel("Date")
+    plt.ylabel("Return")
+    plt.title("Monthly Portfolio Returns Over Time")
+    plt.legend()
+    plt.grid(True)
+    plt.show()
+
+def monthly_alpha():
+    portfolio = portfolio_returns()
+
+    portfolio['TYPE'] = np.where(portfolio['PORTFOLIO'].isin(['LongPosV1', 'LongPosV2']), 'LONG', 'SHORT')
+    alpha = portfolio.groupby(['DATE', 'TYPE'])['RET'].apply(
+        RET=lambda x: x.diff()
+    ).reset_index()
+    portfolio = portfolio.merge(alpha, how='left', on='DATE')
+
+    print(portfolio)
+
+
+monthly_alpha()

### 2025-02-27 monthly_returns.py
Fixed Cum Returns

@@ -31,9 +31,12 @@ def portfolio_returns():
     L_v2 = monthly_returns(Pos="Long", Port="V2")
     S_v2 = monthly_returns(Pos="Short", Port="V2")
 
+    S_v1['RET'] = S_v1['RET'] * -1
+    S_v2['RET'] = S_v2['RET'] * -1
+
     portfolio = pd.concat([L_v1, S_v1, L_v2, S_v2])
     portfolio = portfolio.groupby(['DATE', 'PORTFOLIO']).agg({'RET': 'mean', 'IVOL': 'mean'}).reset_index()
 
     return portfolio
 
-portfolio_returns()
+print(portfolio_returns())

### 2025-02-27 visuals.py

@@ -37,8 +37,9 @@ def monthly_ivol():
 def monthly_total_return():
 
     portfolio = portfolio_returns()
-    portfolio['CUM_RET'] = (1 + portfolio['RET']).cumprod()
+    portfolio['CUM_RET'] = portfolio.groupby(['PORTFOLIO'])['RET'].transform(lambda x: np.cumprod(1 + x))
 
+    portfolio.to_csv(r"c:\Users\brayd\Downloads\test.csv")
     plt.figure(figsize=(12, 6))
     for port in portfolio['PORTFOLIO'].unique():
         data = portfolio[portfolio['PORTFOLIO'] == port]
@@ -51,16 +52,5 @@ def monthly_total_return():
     plt.grid(True)
     plt.show()
 
-def monthly_alpha():
-    portfolio = portfolio_returns()
 
-    portfolio['TYPE'] = np.where(portfolio['PORTFOLIO'].isin(['LongPosV1', 'LongPosV2']), 'LONG', 'SHORT')
-    alpha = portfolio.groupby(['DATE', 'TYPE'])['RET'].apply(
-        RET=lambda x: x.diff()
-    ).reset_index()
-    portfolio = portfolio.merge(alpha, how='left', on='DATE')
-
-    print(portfolio)
-
-
-monthly_alpha()
+monthly_total_return()

### 2025-02-27 visuals.py
Added Alpha Graphs

@@ -35,11 +35,9 @@ def monthly_ivol():
     plt.show()
 
 def monthly_total_return():
-
     portfolio = portfolio_returns()
     portfolio['CUM_RET'] = portfolio.groupby(['PORTFOLIO'])['RET'].transform(lambda x: np.cumprod(1 + x))
 
-    portfolio.to_csv(r"c:\Users\brayd\Downloads\test.csv")
     plt.figure(figsize=(12, 6))
     for port in portfolio['PORTFOLIO'].unique():
         data = portfolio[portfolio['PORTFOLIO'] == port]
@@ -48,6 +46,28 @@ def monthly_total_return():
     plt.xlabel("Date")
     plt.ylabel("Return")
     plt.title("Monthly Portfolio Returns Over Time")
+    plt.legend()
+    plt.grid(True)
+    plt.show()
+
+def monthly_alpha():
+    portfolio = portfolio_returns()
+    portfolio = portfolio[['DATE', 'PORTFOLIO', 'RET']]
+    portfolio = portfolio.pivot(index='DATE', columns='PORTFOLIO', values='RET')
+
+    portfolio['ALPHA_L'] = portfolio['LongPosV2'] - portfolio['LongPosV1']
+    portfolio['ALPHA_S'] = portfolio['ShortPosV2'] - portfolio['ShortPosV1']
+
+    portfolio['ALPHA_L'] = (1 + portfolio['ALPHA_L']).cumprod()
+    portfolio['ALPHA_S'] = (1 + portfolio['ALPHA_S']).cumprod()
+    
+    plt.figure(figsize=(12, 6))
+    plt.plot(portfolio.index, portfolio['ALPHA_L'], label='Alpha Long')
+    plt.plot(portfolio.index, portfolio['ALPHA_S'], label='Alpha Short')
+
+    plt.xlabel("Date")
+    plt.ylabel("Alpha")
+    plt.title("Monthly Portfolio Alpha Over Time")
     plt.legend()
     plt.grid(True)
     plt.show()

### 2025-03-03 visuals.py
Added Histograms

@@ -3,6 +3,7 @@ from monthly_returns import portfolio_returns
 import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
+from scipy.stats import norm
 
 def monthly_ret():
     portfolio = portfolio_returns()
@@ -72,5 +73,31 @@ def monthly_alpha():
     plt.grid(True)
     plt.show()
 
+def monthly_hist():
+    portfolio = portfolio_returns()
+    unique_ports = portfolio['PORTFOLIO'].unique()
+    num_ports = len(unique_ports)
+    num_rows = num_cols = 2
 
-monthly_total_return()
+    fig, axs = plt.subplots(num_rows, num_cols, figsize=(12, 12))
+    axs = axs.flatten()
+
+    for ax, port in zip(axs, unique_ports):
+        data = portfolio[portfolio['PORTFOLIO'] == port]
+        ax.hist(data['RET'], bins=20, alpha=0.5, label=port, density=True)
+
+        mean, std_dev = norm.fit(data['RET'])
+        
+        x = np.linspace(min(data['RET']), max(data['RET']), 100)
+        p = norm.pdf(x, mean, std_dev)
+        ax.plot(x, p, 'k', linewidth=2, label='Normal dist.')
+
+        ax.set_xlabel("Return")
+        ax.set_ylabel("Frequency")
+        ax.set_title("")
+        ax.grid(True)
+        ax.legend()
+        
+    plt.show()
+
+monthly_hist()

### 2025-03-04 paper.qmd
Added Charts To Paper

--- 2025-02-25 paper.qmd
@@
 title: "The Tail Risk Premium - Extension of Lottery Preference Premium"
-format: html
-
+author: "Brayden Boyko, Roan Henderson, Andrew Pettifor"
+format:
+  html:
+    code_folding: hide
+    self_contained: true
+    toc: false
+    dev: "png"
+    fig_embed: true
+    embed-resources: true
+execute:
+  echo: false
+jupyter:
+  kernel: python3
 ---
@@
 + Adjustments: Based on performance analysis, adjustments may be made to the portfolio composition, selection criteria, or rebalancing frequency to enhance the strategy's overall performance and resilience to market fluctuations.
+
+## Performance Evaluations
+A variety of benchmark statistics will be included in our portfolio testing:
+
++ Annual / Monthly Return Values/Charts For 4 Portfolios
++ Description of Stock Breakdown For Four Portfolios (Skew, IVOL)
++ V1 & V2 Screening
+
+### Monthly Returns
+
+```{python}
+from visuals import monthly_total_return
+
+monthly_total_return()
+
+```
+
+#### Monthly Return Histogram
+```{python}
+from visuals import monthly_hist
+
+monthly_hist()
+```
+
+#### Annual Returns
+
+#### V1 Screen: High-Low IVOL
+
+```{python}
+from visuals import monthly_ivol
+
+monthly_ivol()
+```
+
+#### V2 Screen: Negative-Positive Skew
+
+#### V2 Alpha
+
+```{python}
+from visuals import monthly_alpha
+
+monthly_alpha()
+
+```

### 2025-03-04 monthly_returns.py

@@ -38,5 +38,3 @@ def portfolio_returns():
     portfolio = portfolio.groupby(['DATE', 'PORTFOLIO']).agg({'RET': 'mean', 'IVOL': 'mean'}).reset_index()
 
     return portfolio
-
-print(portfolio_returns())

### 2025-03-04 visuals.py

@@ -8,7 +8,7 @@ from scipy.stats import norm
 def monthly_ret():
     portfolio = portfolio_returns()
 
-    plt.figure(figsize=(12, 6))
+    plt.figure(figsize=(8, 5))
     for port in portfolio['PORTFOLIO'].unique():
         data = portfolio[portfolio['PORTFOLIO'] == port]
         plt.plot(data['DATE'], data['RET'], label=port)
@@ -22,8 +22,9 @@ def monthly_ret():
 
 def monthly_ivol():
     portfolio = portfolio_returns()
-
-    plt.figure(figsize=(12, 6))
+    portfolio = portfolio[portfolio['PORTFOLIO'].isin(['LongPosV1', 'ShortPosV1'])]
+    
+    plt.figure(figsize=(8, 5))
     for port in portfolio['PORTFOLIO'].unique():
         data = portfolio[portfolio['PORTFOLIO'] == port]
         plt.plot(data['DATE'], data['IVOL'], label=port)
@@ -39,7 +40,7 @@ def monthly_total_return():
     portfolio = portfolio_returns()
     portfolio['CUM_RET'] = portfolio.groupby(['PORTFOLIO'])['RET'].transform(lambda x: np.cumprod(1 + x))
 
-    plt.figure(figsize=(12, 6))
+    plt.figure(figsize=(8, 5))
     for port in portfolio['PORTFOLIO'].unique():
         data = portfolio[portfolio['PORTFOLIO'] == port]
         plt.plot(data['DATE'], data['CUM_RET'], label=port)
@@ -53,16 +54,14 @@ def monthly_total_return():
 
 def monthly_alpha():
     portfolio = portfolio_returns()
-    portfolio = portfolio[['DATE', 'PORTFOLIO', 'RET']]
-    portfolio = portfolio.pivot(index='DATE', columns='PORTFOLIO', values='RET')
+    portfolio['CUM_RET'] = portfolio.groupby(['PORTFOLIO'])['RET'].transform(lambda x: np.cumprod(1 + x))
+    portfolio = portfolio[['DATE', 'PORTFOLIO', 'CUM_RET']]
+    portfolio = portfolio.pivot(index='DATE', columns='PORTFOLIO', values='CUM_RET')
 
     portfolio['ALPHA_L'] = portfolio['LongPosV2'] - portfolio['LongPosV1']
[…]

### 2025-03-04 portfolio.py
Fixed Error In Ticker Selection

@@ -4,7 +4,7 @@ import numpy as np
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile2)
+    df = pd.read_csv(dataFile)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)
@@ -85,7 +85,6 @@ def portfolio():
     v1_df = v1_df.pivot(index=['Year', 'TICKER'], columns='Position', values=['T_RETURN', 'SKEW']).reset_index()
     v1_df.columns = ['Year', 'TICKER', 'LongRetV1', 'ShortRetV1', 'LongSkewV1', 'ShortSkewV1']
     v1_df = v1_df[v1_df['Year'] >= 2007] ## Rebalances Begin In 2007
-
     rebalance_years = list(v1_df['Year'].unique())
 
     v2_list = []
@@ -110,29 +109,34 @@ def portfolio():
     ## Long-Short Base Startegy (L_Base/S_Base) ---> Long-Short Portfolio Startegy (L_Port/S_Port)
 
     L_Base = v1_df[['Year', 'TICKER', 'LongRetV1', 'LongSkewV1']].dropna()
+    tickers = L_Base.groupby('Year')['TICKER'].unique().reset_index(name='LongPosV1')
     L_Base = L_Base.groupby('Year')['LongRetV1'].agg(LongV1Ret='mean').reset_index()
-    tickers = v1_df.groupby('Year')['TICKER'].unique().reset_index(name='LongPosV1')
     L_Base = L_Base.merge(tickers, on='Year', how='left')
 
     S_Base = v1_df[['Year', 'TICKER', 'ShortRetV1', 'ShortSkewV1']].dropna()
+    tickers = S_Base.groupby('Year')['TICKER'].unique().reset_index(name='ShortPosV1')
     S_Base = S_Base.groupby('Year')['ShortRetV1'].agg(ShortV1Ret='mean').reset_index()
     S_Base['ShortV1Ret'] = (S_Base['ShortV1Ret'] * -1).clip(lower=-1) ## Make Returns Inverse
-    tickers = v1_df.groupby('Year')['TICKER'].unique().reset_index(name='ShortPosV1')
[…]

### 2025-03-04 visuals.py

@@ -1,4 +1,5 @@
 from monthly_returns import portfolio_returns
+from portfolio import pchart
 
 import pandas as pd
 import numpy as np
@@ -24,16 +25,16 @@ def monthly_ivol():
     portfolio = portfolio_returns()
     portfolio = portfolio[portfolio['PORTFOLIO'].isin(['LongPosV1', 'ShortPosV1'])]
     
-    plt.figure(figsize=(8, 5))
-    for port in portfolio['PORTFOLIO'].unique():
+    fig, axs = plt.subplots(2, 1, figsize=(8, 8))
+    
+    for i, port in enumerate(portfolio['PORTFOLIO'].unique()):
         data = portfolio[portfolio['PORTFOLIO'] == port]
-        plt.plot(data['DATE'], data['IVOL'], label=port)
-
-    plt.xlabel("Date")
-    plt.ylabel("IVOL")
-    plt.title("Monthly Portfolio Idiosyncratic Volatility Over Time")
-    plt.legend()
-    plt.grid(True)
+        axs[i].plot(data['DATE'], data['IVOL'], label=port)
+        axs[i].set_xlabel("Date")
+        axs[i].set_ylabel("IVOL")
+        axs[i].legend()
+        axs[i].grid(True)
+    
     plt.show()
 
 def monthly_total_return():
@@ -96,5 +97,26 @@ def monthly_hist():
         ax.set_title("")
         ax.grid(True)
         ax.legend()
+
+    plt.show()
+
+
+def annual_ret():
+    data = pchart()
+    data = data[['Year', 'LongV1Ret', 'LongV2Ret', 'ShortV1Ret', 'ShortV2Ret']]
+    data = data.melt(id_vars=['Year'], var_name='PORTFOLIO', value_name='RET')
+
+    fig, axs = plt.subplots(2, 2, figsize=(8, 8), sharex=True)
+
+    portfolios = ['LongV1Ret', 'LongV2Ret', 'ShortV1Ret', 'ShortV2Ret']
+    titles = ['Long V1 Returns', 'Long V2 Returns', 'Short V1 Returns', 'Short V2 Returns']
+
+    for ax, portfolio, title in zip(axs.flatten(), portfolios, titles):
+        subset = data[data['PORTFOLIO'] == portfolio]
+        ax.bar(subset['Year'], subset['RET'], color='teal')
+        ax.set_title(title)
+        ax.set_ylabel('Return')
[…]

### 2025-03-04 portfolio.py
Updated Report Structure

@@ -94,12 +94,12 @@ def portfolio():
         v2_long = v2_df[['Year', 'TICKER','LongRetV1', 'LongSkewV1']].dropna()
         v2_long.columns = ['Year', 'TICKER', 'LongRetV2', 'LongSkewV2']
         v2_long = v2_long.sort_values(by='LongSkewV2', ascending=True)
-        v2_long = v2_long.tail(30)
+        v2_long = v2_long.tail(30) ## OG STRATEGY: v2_long.tail(30)
 
         v2_short = v2_df[['Year', 'TICKER','ShortRetV1', 'ShortSkewV1']].dropna()
         v2_short.columns = ['Year', 'TICKER', 'ShortRetV2', 'ShortSkewV2']
         v2_short = v2_short.sort_values(by='ShortSkewV2', ascending=True)
-        v2_short = v2_short.head(30)
+        v2_short = v2_short.head(30) ## OG STRATEGY: v2_long.head(30)
 
         combined = pd.concat([v2_long, v2_short], axis=0)
         v2_list.append(combined)

### 2025-03-04 portfolio.py
Added Ticker List

@@ -4,7 +4,7 @@ import numpy as np
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile)
+    df = pd.read_csv(dataFile2)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)
@@ -133,8 +133,6 @@ def portfolio():
     port = L_Base.merge(L_Port, how='left', on='Year')
     port = port.merge(S_Base, how='left', on='Year')
     port = port.merge(S_Port, how='left', on='Year')
-
-    port.to_csv(r"C:\Users\Brayden Boyko\Downloads\tester.csv")
 
     return port
 

### 2025-03-04 sumStats.py

@@ -1,0 +1,18 @@
+from portfolio import pchart
+
+import pandas as pd
+
+def tickers():
+
+    data = pchart()
+
+    data = data[['Year', 'LongPosV2', 'ShortPosV2']]
+    exploded_long = data.explode('LongPosV2').rename(columns={'LongPosV2': 'Ticker'}).assign(Portfolio='Long')
+    exploded_short = data.explode('ShortPosV2').rename(columns={'ShortPosV2': 'Ticker'}).assign(Portfolio='Short')
+
+    data = pd.concat([exploded_long, exploded_short], ignore_index=True)
+    data = data[['Year', 'Portfolio', 'Ticker']]
+
+    return data
+
+tickers()

### 2025-03-04 visuals.py

@@ -1,10 +1,12 @@
 from monthly_returns import portfolio_returns
 from portfolio import pchart
+from sumStats import tickers
 
 import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
 from scipy.stats import norm
+from wordcloud import WordCloud
 
 def monthly_ret():
     portfolio = portfolio_returns()
@@ -119,4 +121,27 @@ def annual_ret():
         ax.set_xlabel('Year')
         ax.tick_params(axis='x', rotation=45)
 
+    plt.show()
+
+def ticker_count():
+    data = tickers()
+    data['Position'] = data.groupby(['Year', 'Portfolio']).cumcount() + 1
+    data['Portfolio_Value'] = data['Portfolio'].map({'Long': 1, 'Short': -1})
+    data['plot'] = data['Position'] * data['Portfolio_Value']
+    
+    plt.figure(figsize=(8, 8))
+
+    for _, row in data.iterrows():
+        plt.scatter(row['Year'], row['plot'], s=100, 
+                    color='white' if row['Portfolio'] == 'Long' else 'white')
+
+        plt.text(row['Year'], row['plot'], row['Ticker'], 
+                 ha='center', va='center', fontsize=5, rotation=0)
+
+    plt.axhline(0, color='black', linewidth=1) 
+    plt.xticks(sorted(data['Year'].unique()))
+    plt.yticks([])
+    plt.xlabel("Year")
+    plt.ylabel("Short (Bottom) vs Long (Top)")
+    plt.title("Ticker Portfolio Positions by Year")
     plt.show()

### 2025-03-06 portfolio.py
Updates to SKEW calc

@@ -45,16 +45,20 @@ def portfolio():
 
     ## V1->V2 Screen (Extreme IVOL High-Low) -> (Extreme SKEW Positive-Negative)
     for period in rebalance_period:
-        df_back = df[(df['YEAR'] <= (period - 1)) & (df['YEAR'] >= (period - 6))].copy()  ## Five Year Lookback Window
+        df_back = df[(df['YEAR'] <= (period - 1)) & (df['YEAR'] >= (period - 4))].copy()  ## Five Year Lookback Window
         df_back['PERIOD'] = period ## Rebalance Year
         
         stats_df = df_back.groupby('TICKER')['IVOL'].agg(
             AVE='mean',
             ).reset_index()
         
-        stats2_df = df_back.groupby('TICKER')['RET'].agg(
-            SKEW='skew'
+        stats2_df = df_back.groupby('TICKER').agg(
+            SKEW=('RET','skew'),
+            RET_M =('RET', 'mean'),
+            AVE_I=('IVOL', 'mean')
         ).reset_index()
+
+        stats2_df['SKEW'] = stats2_df['SKEW_ADJ'] = stats2_df['SKEW'] * stats2_df['RET_M'].apply(lambda x: -abs(x)) / stats2_df['AVE_I'] ## THIS IS OUR FILTER
         
         return_df = df[df['YEAR'] == period]
         return_df = df_back.groupby('TICKER')['RET'].agg(
@@ -94,12 +98,12 @@ def portfolio():
         v2_long = v2_df[['Year', 'TICKER','LongRetV1', 'LongSkewV1']].dropna()
         v2_long.columns = ['Year', 'TICKER', 'LongRetV2', 'LongSkewV2']
         v2_long = v2_long.sort_values(by='LongSkewV2', ascending=True)
-        v2_long = v2_long.tail(30) ## OG STRATEGY: v2_long.tail(30)
+        v2_long = v2_long.nlargest(30, 'LongSkewV2') ## OG STRATEGY: v2_long.tail(30)
 
         v2_short = v2_df[['Year', 'TICKER','ShortRetV1', 'ShortSkewV1']].dropna()
         v2_short.columns = ['Year', 'TICKER', 'ShortRetV2', 'ShortSkewV2']
         v2_short = v2_short.sort_values(by='ShortSkewV2', ascending=True)
-        v2_short = v2_short.head(30) ## OG STRATEGY: v2_long.head(30)
[…]

### 2025-03-09 portfolio.py
Updates To Annual Graphics

@@ -4,7 +4,7 @@ import numpy as np
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile2)
+    df = pd.read_csv(dataFile)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)
@@ -61,9 +61,8 @@ def portfolio():
         stats2_df['SKEW'] = stats2_df['SKEW_ADJ'] = stats2_df['SKEW'] * stats2_df['RET_M'].apply(lambda x: -abs(x)) / stats2_df['AVE_I'] ## THIS IS OUR FILTER
         
         return_df = df[df['YEAR'] == period]
-        return_df = df_back.groupby('TICKER')['RET'].agg(
-            T_RETURN='sum'
-            ).reset_index()
+        return_df = df_back.groupby('TICKER')['RET'].apply(lambda x: (x + 1).prod() - 1).reset_index(name='T_RETURN')
+
         
         df_back = df_back.merge(stats_df, on='TICKER', how='left')
         df_back = df_back.merge(stats2_df, on='TICKER', how='left')
@@ -116,6 +115,7 @@ def portfolio():
     tickers = L_Base.groupby('Year')['TICKER'].unique().reset_index(name='LongPosV1')
     L_Base = L_Base.groupby('Year')['LongRetV1'].agg(LongV1Ret='mean').reset_index()
     L_Base = L_Base.merge(tickers, on='Year', how='left')
+    
 
     S_Base = v1_df[['Year', 'TICKER', 'ShortRetV1', 'ShortSkewV1']].dropna()
     tickers = S_Base.groupby('Year')['TICKER'].unique().reset_index(name='ShortPosV1')
@@ -142,3 +142,4 @@ def portfolio():
 
 def pchart():
     return portfolio()
+

### 2025-03-09 visuals.py

@@ -6,7 +6,6 @@ import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
 from scipy.stats import norm
-from wordcloud import WordCloud
 
 def monthly_ret():
     portfolio = portfolio_returns()
@@ -104,24 +103,39 @@ def monthly_hist():
 
 
 def annual_ret():
-    data = pchart()
-    data = data[['Year', 'LongV1Ret', 'LongV2Ret', 'ShortV1Ret', 'ShortV2Ret']]
-    data = data.melt(id_vars=['Year'], var_name='PORTFOLIO', value_name='RET')
+    portfolio = portfolio_returns()
+    portfolio['CUM_RET'] = portfolio.groupby(['PORTFOLIO'])['RET'].transform(lambda x: np.cumprod(1 + x))
+    portfolio['MONTH'] = portfolio['DATE'].dt.month
+    portfolio['YEAR'] = portfolio['DATE'].dt.year
+    portfolio = portfolio[portfolio['MONTH'] == 12]
+    portfolio['INVESTMENT'] = portfolio['CUM_RET'] * 100_000
 
-    fig, axs = plt.subplots(2, 2, figsize=(8, 8), sharex=True)
+    portfolio_v1 = portfolio[portfolio['PORTFOLIO'].str.contains('Long')]
+    portfolio_v2 = portfolio[portfolio['PORTFOLIO'].str.contains('Short')]
 
-    portfolios = ['LongV1Ret', 'LongV2Ret', 'ShortV1Ret', 'ShortV2Ret']
-    titles = ['Long V1 Returns', 'Long V2 Returns', 'Short V1 Returns', 'Short V2 Returns']
+    pivot_v1 = portfolio_v1.pivot(index='YEAR', columns='PORTFOLIO', values='INVESTMENT')
+    pivot_v2 = portfolio_v2.pivot(index='YEAR', columns='PORTFOLIO', values='INVESTMENT')
 
-    for ax, portfolio, title in zip(axs.flatten(), portfolios, titles):
-        subset = data[data['PORTFOLIO'] == portfolio]
-        ax.bar(subset['Year'], subset['RET'], color='teal')
+    fig, axes = plt.subplots(2, 1, figsize=(8, 8), sharey=True)
+    width = 0.4 ## BAR WIDTH
+
+    for ax, pivot, title in zip(axes, [pivot_v1, pivot_v2], ["Long Portfolios", "Short Portfolios"]):
+        years = pivot.index
+        portfolios = pivot.columns
+
[…]

### 2025-03-09 visuals.py
Added total equal-weighted portfolios

@@ -159,3 +159,30 @@ def ticker_count():
     plt.ylabel("Short (Bottom) vs Long (Top)")
     plt.title("Ticker Portfolio Positions by Year")
     plt.show()
+
+def portfolio_totals():
+    portfolio = portfolio_returns()
+    portfolio['CUM_RET'] = portfolio.groupby(['PORTFOLIO'])['RET'].transform(lambda x: np.cumprod(1 + x))
+    v1 = portfolio[portfolio['PORTFOLIO'].str.contains('V1')]
+    v2 = portfolio[portfolio['PORTFOLIO'].str.contains('V2')]
+    
+    ## EQUAL WEIGHT
+    v1 = v1.groupby('DATE')['CUM_RET'].agg(V1='mean').reset_index()
+    v2 = v2.groupby('DATE')['CUM_RET'].agg(V2='mean').reset_index()
+
+    port = v1.merge(v2, how='left', on='DATE')
+
+    plt.figure(figsize=(8, 5))
+    plt.plot(port['DATE'], port['V1'], label='Base Portfolio')
+    plt.plot(port['DATE'], port['V2'], label='Strategy Portfolio')
+
+    plt.xlabel("Date")
+    plt.ylabel("Returns")
+    plt.title("Total Equal-Weighted Portfolio Returns Over Time")
+    plt.legend()
+    plt.grid(True)
+    plt.show()
+
+
+
+

### 2025-03-11 paper.qmd
Fixed Error In Ticker Selection
Added Ticker List
Updates To Annual Graphics
Added total equal-weighted portfolios
Added V2 Screening

@@
 
-#### Annual Returns
+#### Annual Returns (Value of $100k Invested)
+```{python}
+from visuals import annual_ret
+annual_ret()
+
+```
 
@@
 
+```{python}
+from visuals import portfolio_skews
+
+portfolio_skews()
+```
+
 #### V2 Alpha
@@
 ```
+
+#### V2 Portfolio Positions Over Time
+
+```{python}
+from visuals import ticker_count
+
+ticker_count()
+```
+
+#### Equal Weighted Portfolios (Baseline vs Strategy-Extension)
+
+```{python}
+from visuals import portfolio_totals
+
+portfolio_totals()
+```

### 2025-03-11 portfolio.py

@@ -113,25 +113,24 @@ def portfolio():
 
     L_Base = v1_df[['Year', 'TICKER', 'LongRetV1', 'LongSkewV1']].dropna()
     tickers = L_Base.groupby('Year')['TICKER'].unique().reset_index(name='LongPosV1')
-    L_Base = L_Base.groupby('Year')['LongRetV1'].agg(LongV1Ret='mean').reset_index()
+    L_Base = L_Base.groupby('Year').agg(LongV1Ret=('LongRetV1', 'mean'), LongV1Skew=('LongSkewV1', 'mean')).reset_index()
     L_Base = L_Base.merge(tickers, on='Year', how='left')
-    
 
     S_Base = v1_df[['Year', 'TICKER', 'ShortRetV1', 'ShortSkewV1']].dropna()
     tickers = S_Base.groupby('Year')['TICKER'].unique().reset_index(name='ShortPosV1')
-    S_Base = S_Base.groupby('Year')['ShortRetV1'].agg(ShortV1Ret='mean').reset_index()
-    S_Base['ShortV1Ret'] = (S_Base['ShortV1Ret'] * -1).clip(lower=-1) ## Make Returns Inverse
+    S_Base = S_Base.groupby('Year').agg(ShortV1Ret=('ShortRetV1', 'mean'), ShortV1Skew=('ShortSkewV1', 'mean')).reset_index()
+    S_Base['ShortV1Ret'] = (S_Base['ShortV1Ret'] * -1).clip(lower=-1)  # Make Returns Inverse
     S_Base = S_Base.merge(tickers, on='Year', how='left')
 
     L_Port = v2_df[['Year', 'TICKER', 'LongRetV2', 'LongSkewV2']].dropna()
     tickers = L_Port.groupby('Year')['TICKER'].unique().reset_index(name='LongPosV2')
-    L_Port = L_Port.groupby('Year')['LongRetV2'].agg(LongV2Ret='mean').reset_index()
+    L_Port = L_Port.groupby('Year').agg(LongV2Ret=('LongRetV2', 'mean'), LongV2Skew=('LongSkewV2', 'mean')).reset_index()
     L_Port = L_Port.merge(tickers, on='Year', how='left')
 
     S_Port = v2_df[['Year', 'TICKER', 'ShortRetV2', 'ShortSkewV2']].dropna()
     tickers = S_Port.groupby('Year')['TICKER'].unique().reset_index(name='ShortPosV2')
[…]

### 2025-03-11 visuals.py

@@ -183,6 +183,31 @@ def portfolio_totals():
     plt.grid(True)
     plt.show()
 
+    
+def portfolio_skews():
+    portfolio = pchart()
+    port = portfolio[['Year', 'LongV1Skew', 'ShortV1Skew', 'LongV2Skew', 'ShortV2Skew']]
+
+    fig, axes = plt.subplots(2, 2, figsize=(8, 8), sharex=True)
+
+    skews = {
+        "LongV1Skew": (0, 0),
+        "ShortV1Skew": (0, 1),
+        "LongV2Skew": (1, 0),
+        "ShortV2Skew": (1, 1)
+    }
+
+    for key, (row, col) in skews.items():
+        ax = axes[row, col]
+        ax.plot(port["Year"], port[key], marker="o", linestyle="-", label=key)
+        ax.set_title(key)
+        ax.set_xlabel("Year ")
+        ax.set_ylabel("Skew")
+        ax.legend()
+        ax.grid(True)
+
+    plt.tight_layout()
+    plt.show()
 
 
 

### 2025-03-16 FIN 416 Tables.qmd
Add files via upload

---
title: "FIN 416 Tables"
author: "Andrew Pettifor"
format: html
editor: visual
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(tidyr)
library(tidyselect)
library(stats)
library(PerformanceAnalytics)
library(xts)
library(gt)
library(RTL)
library(plotly)
library(ggplot2)
library(arrow)
library(patchwork)
library(tidyquant)
library(scales)
library(ggridges)
library(viridis)
library(forcats)
library(scales)
library(lubridate)
library(stats)
library(broom)
library(gt)
library(car)
library(GGally)
library(rlang)
library(glue)
library(purrr)
library(ggfortify)
library(ggpmisc)

```

```{r echo=FALSE}

original_df <- read.csv("https://raw.githubusercontent.com/boykowealth/TailRiskPremium_PortfolioConcept/refs/heads/main/RetVolSkew_Monthly.csv")

port_df_long <- original_df %>% 
  select(DATE, PORTFOLIO, RET, IVOL, SKEW)

port_df_wide <- port_df_long %>% 
  pivot_wider(names_from = PORTFOLIO, values_from = c(RET,IVOL,SKEW))

port_returns_wide <- port_df_wide %>% 
  select(DATE,RET_LongPosV1,RET_LongPosV2,RET_ShortPosV1,RET_ShortPosV2) %>% 
  mutate(RET_V1_Combined = RET_LongPosV1+RET_ShortPosV1) %>% 
  mutate(RET_V2_Combined = RET_LongPosV2 + RET_ShortPosV2)

return_columns <- grep("RET", names(port_returns_wide), value = TRUE)

port_returns_annualized <- port_returns_wide %>%
  mutate(across(all_of(return_columns), ~ (1 + .)^12 - 1))

```

# Slide 14 Performance Evaluation: New vs. Base Overall Strategy Return Summary

Need: Mean, St.Dev, Max number of Observations, T Statistic, Sharpe Ratio, Correlation

```{r echo=FALSE}

slide14_data <- port_returns_annualized %>% 
  select(DATE,RET_V1_Combined, RET_V2_Combined) %>% 
  mutate(Base = RET_V1_Combined,
         New = RET_V2_Combined,
         NewSubBase = RET_V2_Combined-RET_V1_Combined)

slide14_table <- slide14_data %>% 
  summarise(
    across(c(Base, New, NewSubBase), list(
      Mean = ~ mean(.x, na.rm = TRUE),
      StDev = ~ sd(.x, na.rm = TRUE),
      Max = ~ max(.x, na.rm = TRUE),
      Obs = ~ sum(!is.na(.x)),
      TStat = ~ mean(.x, na.rm = TRUE) / (sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x)))),
      Sharpe = ~ mean(.x, na.rm = TRUE) / sd(.x, na.rm = TRUE)
    )),
    Correlation = cor(Base, New, use = "complete.obs")
  )

summary_table_14 <- slide14_table %>%
  pivot_longer(everything(), names_to = c("Column", "Metric"), names_sep = "_", values_to = "Value") %>%
  pivot_wider(names_from = Column, values_from = Value)

summary_table_14 %>% gt()

```

# Slide 15 New vs. Base Overall Strategy Return Dynamics

```{r echo=FALSE}

Slide15_Data <- slide14_data %>% 
  select(DATE, Base, New) %>% 
  pivot_longer(cols = c(Base, New),
               names_to = "Portfolio",
               values_to = "Returns")

ggplot(Slide15_Data, aes(x = as.Date(DATE), y = Returns, color = Portfolio, group = Portfolio)) +
  geom_line(size = 1) + 
  labs(title = "Base vs. New Portfolio Returns Over Time",
       subtitle = "Annualized Monthly Returns",
       x = "Date",
       y = "Returns %",
       color = "Portfolio") +
  theme_minimal() +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 year") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +  # Converts y-axis to percentage

  theme(legend.position = "top",
  axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

# Slide 17 New vs Base Long
```{r echo=FALSE}

slide17_data <- port_returns_annualized %>% 
  select(DATE,RET_LongPosV1, RET_LongPosV2) %>% 
  mutate(BaseLong = RET_LongPosV1,
         NewLong = RET_LongPosV2,
         NewSubBaseLong = RET_LongPosV2-RET_LongPosV1)

slide17_table <- slide17_data %>% 
  summarise(
    across(c(BaseLong, NewLong, NewSubBaseLong), list(
      Mean = ~ mean(.x, na.rm = TRUE),
      StDev = ~ sd(.x, na.rm = TRUE),
      Max = ~ max(.x, na.rm = TRUE),
      Obs = ~ sum(!is.na(.x)),
      TStat = ~ mean(.x, na.rm = TRUE) / (sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x)))),
      Sharpe = ~ mean(.x, na.rm = TRUE) / sd(.x, na.rm = TRUE)
    )),
    Correlation = cor(BaseLong, NewLong, use = "complete.obs")
  )

summary_table_17 <- slide17_table %>%
  pivot_longer(everything(), names_to = c("Column", "Metric"), names_sep = "_", values_to = "Value") %>%
  pivot_wider(names_from = Column, values_from = Value)

summary_table_17 %>% gt()

```

# Slide 17 Visual
```{r echo=FALSE}

Slide17_Data <- slide17_data %>% 
  select(DATE, BaseLong, NewLong) %>% 
  pivot_longer(cols = c(BaseLong, NewLong),
               names_to = "Portfolio",
               values_to = "Returns")

ggplot(Slide17_Data, aes(x = as.Date(DATE), y = Returns, color = Portfolio, group = Portfolio)) +
  geom_line(size = 1) + 
  labs(title = "Base Long vs. New Long Portfolio Returns Over Time",
       subtitle = "Annualized Monthly Returns",
       x = "Date",
       y = "Returns %",
       color = "Portfolio") +
  theme_minimal() +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 year") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +

  theme(legend.position = "top",
  axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

# Slide 18: New vs. Base Short Side Return Summary Statistics & Hypothesis Test

```{r echo=FALSE}
slide18_data <- port_returns_annualized %>% 
  select(DATE,RET_ShortPosV1, RET_ShortPosV2) %>% 
  mutate(BaseShort = RET_ShortPosV1,
         NewShort = RET_ShortPosV2,
         NewSubBaseShort = RET_ShortPosV2-RET_ShortPosV1)

slide18_table <- slide18_data %>% 
  summarise(
    across(c(BaseShort, NewShort, NewSubBaseShort), list(
      Mean = ~ mean(.x, na.rm = TRUE),
      StDev = ~ sd(.x, na.rm = TRUE),
      Max = ~ max(.x, na.rm = TRUE),
      Obs = ~ sum(!is.na(.x)),
      TStat = ~ mean(.x, na.rm = TRUE) / (sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x)))),
      Sharpe = ~ mean(.x, na.rm = TRUE) / sd(.x, na.rm = TRUE)
    )),
    Correlation = cor(BaseShort, NewShort, use = "complete.obs")
  )

summary_table_18 <- slide18_table %>%
  pivot_longer(everything(), names_to = c("Column", "Metric"), names_sep = "_", values_to = "Value") %>%
  pivot_wider(names_from = Column, values_from = Value)

summary_table_18 %>% gt()

```

# Slide 18 Visual
```{r echo=FALSE}

Slide18_Data <- slide18_data %>% 
  select(DATE, BaseShort, NewShort) %>% 
  pivot_longer(cols = c(BaseShort, NewShort),
               names_to = "Portfolio",
               values_to = "Returns")

ggplot(Slide18_Data, aes(x = as.Date(DATE), y = Returns, color = Portfolio, group = Portfolio)) +
  geom_line(size = 1) + 
  labs(title = "Base Short vs. New Short Portfolio Returns Over Time",
       subtitle = "Annualized Monthly Returns",
       x = "Date",
       y = "Returns %",
       color = "Portfolio") +
  theme_minimal() +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 year") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +

  theme(legend.position = "top",
  axis.text.x = element_text(angle = 45, hjust = 1)
  )

```
# Slide 21 Table (Im Confused)
```{r echo=FALSE}
slide21_data <- port_returns_annualized %>% 
  select(DATE,RET_LongPosV1,RET_LongPosV2,RET_ShortPosV1,RET_ShortPosV2) %>% 
  mutate(
    BaseLong = RET_LongPosV1,
    BaseShort = RET_ShortPosV1,
    NewLong = RET_LongPosV2,
    NewShort = RET_ShortPosV2
             )

slide21_table <- slide21_data %>% 
  summarise(
    across(c(BaseLong, BaseShort, NewLong, NewShort), list(
      Mean = ~ mean(.x, na.rm = TRUE),
      StDev = ~ sd(.x, na.rm = TRUE),
      Min = ~ min(.x, na.rm = TRUE),
      Max = ~ max(.x, na.rm = TRUE))))

summary_table_21 <- slide21_table %>%
  pivot_longer(everything(), names_to = c("Column", "Metric"), names_sep = "_", values_to = "Value") %>%
  pivot_wider(names_from = Column, values_from = Value)

summary_table_21 %>% gt()
```

### 2025-03-18 paper.qmd
Added Updates To Formatting

@@
 
-### Monthly Returns
+### Four Portfolio Metrics
+
+Four portfolio's were constructed under the assumption of equal weighting dynamics.
+V1 portfolios are comprised of 100 companies and screen through IVOL.
+V2 portfolios describe our teams extension of the V1 sceen, which utilizes skew to screen for 30 companies in each portfolio.
+
+#### Monthly Returns
 
 ```{python}
-from visuals import monthly_total_return
+from visuals import monthly_ret
 
-monthly_total_return()
+monthly_ret()
 
@@
 #### Monthly Return Histogram
+
 ```{python}
@@
 
-#### Annual Returns (Value of $100k Invested)
+#### Cumalitve Monthly Returns (Value of $1 Invested)
+
+```{python}
+from visuals import monthly_total_return
+
+monthly_total_return()
+```
+
+#### Annual Returns (Value of $1 Invested)
 ```{python}
@@
 ```
+
+### Screening Variables
+
+The screening variables are broken into four categories, under the asumption of a three year lookback period.
+Base Long utilizes the lowest 100 IVOL from the dataset each year.
+In contrast, the Base Short utilizes the highest 100 IVOL from the dataset.
 
@@
 
-#### V2 Alpha
-
-```{python}
-from visuals import monthly_alpha
-
-monthly_alpha()
-
-```
+### Portfolio Positions
+The portfolios are equal weighted. This provides an accurate representation of both the V1 and V2 screening effectiveness.
 
@@
 
-#### Equal Weighted Portfolios (Baseline vs Strategy-Extension)
+#### Equal Weighted Portfolios (Baseline vs Strategy-Extension) Cumalitve Returns (Value of $1 Invested)
 
@@
 ```
+
+#### V2 Strategy Alpha Relative To V1 Baseline
+
+```{python}
+from visuals import monthly_alpha
+
+monthly_alpha()
+
+```

### 2025-03-18 monthly_returns.py

@@ -12,15 +12,22 @@ def monthly_returns(Pos=None, Port=None):
     df = data()
     port = portfolio()
     select = f"{Pos}Pos{Port}"
+    skew = f"{Pos}{Port}Skew"
     
-    port = port[['Year', f'{select}']]
+    port = port[['Year', f'{select}', f'{skew}']]
 
     df = df.merge(port, how='left', left_on='YEAR', right_on='Year').dropna()
     df = df.explode(f'{select}')
     df = df[df['TICKER'] == df[f'{select}']]
     df['PORTFOLIO'] = select
 
-    df = df[['DATE', 'YEAR', 'TICKER', 'RET', 'IVOL', 'PORTFOLIO']]
+    ## Converts SKEW Column Headers
+    possible_names = {"ShortV2Skew", "ShortV1Skew", "LongV2Skew", "LongV1Skew"}
+    existing_name = next((col for col in df.columns if col in possible_names), None)
+    if existing_name:
+        df = df.rename(columns={existing_name: "SKEW"})
+
+    df = df[['DATE', 'YEAR', 'TICKER', 'RET', 'IVOL', 'SKEW', 'PORTFOLIO']]
 
     return df
 
@@ -35,6 +42,6 @@ def portfolio_returns():
     S_v2['RET'] = S_v2['RET'] * -1
 
     portfolio = pd.concat([L_v1, S_v1, L_v2, S_v2])
-    portfolio = portfolio.groupby(['DATE', 'PORTFOLIO']).agg({'RET': 'mean', 'IVOL': 'mean'}).reset_index()
+    portfolio = portfolio.groupby(['DATE', 'PORTFOLIO']).agg({'RET': 'mean', 'IVOL': 'mean', 'SKEW': 'mean'}).reset_index()
 
     return portfolio

### 2025-03-18 visuals.py

@@ -5,21 +5,28 @@ from sumStats import tickers
 import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib.ticker as mtick
 from scipy.stats import norm
 
 def monthly_ret():
     portfolio = portfolio_returns()
-
-    plt.figure(figsize=(8, 5))
-    for port in portfolio['PORTFOLIO'].unique():
+    unique_ports = portfolio['PORTFOLIO'].unique()
+    
+    fig, axes = plt.subplots(2, 2, figsize=(8, 8))
+    axes = axes.flatten() 
+    for i, port in enumerate(unique_ports[:4]):  
         data = portfolio[portfolio['PORTFOLIO'] == port]
-        plt.plot(data['DATE'], data['RET'], label=port)
+        ax = axes[i]
+        ax.plot(data['DATE'], data['RET'], label=port)
+        ax.set_xlabel("Date")
+        ax.set_ylabel("Return")
+        ax.set_title(f"Portfolio: {port}")
+        ax.yaxis.set_major_formatter(mtick.PercentFormatter(1.0))
+        ax.legend()
+        ax.grid(True)
 
-    plt.xlabel("Date")
-    plt.ylabel("Return")
-    plt.title("Monthly Portfolio Returns Over Time")
-    plt.legend()
-    plt.grid(True)
+
+    plt.tight_layout()  # Adjust spacing between plots
     plt.show()
 
 def monthly_ivol():
@@ -62,10 +69,11 @@ def monthly_alpha():
 
     portfolio['ALPHA_L'] = portfolio['LongPosV2'] - portfolio['LongPosV1']
     portfolio['ALPHA_S'] = portfolio['ShortPosV2'] - portfolio['ShortPosV1']
+
+    portfolio['ALPHA'] = portfolio['ALPHA_L'] + portfolio['ALPHA_S']
     
     plt.figure(figsize=(8, 5))
-    plt.plot(portfolio.index, portfolio['ALPHA_L'], label='Alpha Long')
-    plt.plot(portfolio.index, portfolio['ALPHA_S'], label='Alpha Short')
+    plt.plot(portfolio.index, portfolio['ALPHA'], label='Portfolio Alpha')
 
     plt.xlabel("Date")
     plt.ylabel("Alpha")
@@ -108,7 +116,7 @@ def annual_ret():
     portfolio['MONTH'] = portfolio['DATE'].dt.month
[…]

### 2025-03-18 FIN 416 Tables.qmd
Added Information Regarding Additional Test

@@
 library(scales)
-library(ggridges)
-library(viridis)
+#library(ggridges)
+#library(viridis)
 library(forcats)
@@
 library(gt)
-library(car)
+#library(car)
 library(GGally)

### 2025-03-18 paper.qmd
Added Information Regarding Additional Test

@@
 ```
+
+### Additional Tests
+
+A series of additional tests were completed to better explain the results found in our data sample period. These tests were conducted in order to improve upon the strategy.
+
+#### Explantion For V2 Short Measure Underperformance
+
+#### Explanation For IVOL Stabalization Post 2008
+
+__Key Takeaways for Post-2008 Market Stability__
+
++ Systemic risk increased post-2008, making diversification ineffective.
++ Correlations tend to rise before a crisis, suggesting a leading indicator for instability.
++ Investor behavior shifted towards liquidity and safety, aligning with our discussion on idiosyncratic risk vs. systemic risk.
++ Market-wide herding intensified, driving the preference for safe assets and reducing the appeal of high-idiosyncratic-risk stocks.
++ Relative decline in idiosyncratic risk due to rising market and industry volatility.
++ Crisis events (2008, COVID-19) still caused firm-specific volatility spikes but within an environment of elevated systematic risk.
++ Investor preference for stability (CSP, large firms) helped reduce firm-specific risk, supporting our discussion on flight to safety and liquidity-driven volatility shifts.
+
+1. "Temporal Evolution of Financial Market Correlations" (Fenn et al.)
+
+__Summary:__ This study applies Principal Component Analysis (PCA) and Random Matrix Theory (RMT) to analyze evolving correlations in financial markets from 1999–2010.
+Key findings indicate that post-2008, financial assets across different markets exhibited much stronger correlations.
+The first principal component accounted for nearly 40% of market variance post-Lehman collapse, suggesting that most assets were moving together.
+This heightened correlation reduces diversification benefits and increases systemic risk.
+
+__Relevance to Post-2008 Stability:__ This aligns with our discussion on correlation risk, where increased asset co-movement reduces the effectiveness of diversification.
+It also supports the high systemic risk in a post-2008 world, showing why investor behavior shifted towards liquidity preference and safe-haven assets.
+
+2. "Correlation of Financial Markets in Times of Crisis" (Sandoval & Franca)
+
+__Summary:__ The study uses eigenvalue decomposition of correlation matrices to show that market correlations increase dramatically during crises (1987, 1998, 2001, 2008).
+Their key finding is that high volatility leads to strong correlations, meaning that during crashes, markets behave as one.
+The study also finds that average market correlation increases prior to a crisis, suggesting an early warning signal.
+
+__Relevance to Post-2008 Stability:__ This reinforces our point that in crisis conditions, herding behavior dominates.
+The finding that correlations increase before crises aligns with the liquidity crunch hypothesis, where investors pull out of riskier assets preemptively.
+The study also supports the behavioral premium idea, where investors chase "lottery stocks" in stable times but rush to liquidity in crises.
+
+3. "Stock Markets Correlation - Before and During the Crisis Analysis" (Moldovan)​"
+
+__Summary:__ This paper specifically analyzes the correlations between New York (Dow Jones), London (FTSE), and Tokyo (Nikkei) before (2003-2006) and during (2007-2010) the crisis.
+It finds that correlations were significantly stronger during the crisis period, meaning that stocks moved together more in downturns than in uptrends.
+Regression models showed that FTSE’s influence on Dow and Nikkei increased during the crisis, while Nikkei and Dow became negatively correlated.
+
+__Relevance to Post-2008 Stability:__ This supports our discussion on global financial interdependence post-2008.
+The increased correlation in downturns confirms that diversification failed when most needed.
+The negative Dow-Nikkei correlation is an interesting insight, possibly reflecting a flight to safety in Japanese markets during the US downturn.

### 2025-03-20 416 Tables, Nearly done.qmd
Added Updated Files

--- 2025-03-16 FIN 416 Tables.qmd
@@
 library(ggpmisc)
+library(devtools)
 
@@
 
-port_returns_annualized <- port_returns_wide %>%
-  mutate(across(all_of(return_columns), ~ (1 + .)^12 - 1))
+port_returns_annualized <- port_returns_wide
 
@@
 
-summary_table_14 %>% gt()
+summary_table_14 %>% 
+   gt() %>% 
+  fmt_percent(
+    columns = where(is.numeric), # Apply only to numeric columns
+    rows = Metric %in% c("Mean", "StDev", "Max", "Correlation"), # Filter only Mean, StDev, and Max
+    decimals = 2
+  ) %>% 
+  fmt_number(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Sharpe", "TStat"), 
+    decimals = 3 # Adjust to desired decimal places
+  ) %>% 
+  fmt_number(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Obs"), 
+    decimals = 0 # Adjust to desired decimal places
+  )
 
@@
 # Slide 17 New vs Base Long
+
 ```{r echo=FALSE}
@@
 
-summary_table_17 %>% gt()
+summary_table_17 %>% 
+  gt() %>% 
+  fmt_percent(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Mean", "StDev", "Max", "Correlation"), 
+    decimals = 2
+  ) %>% 
+  fmt_number(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Sharpe", "TStat"), 
+    decimals = 3 
+  ) %>% 
+  fmt_number(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Obs"), 
+    decimals = 0 
+  )
 
@@
 # Slide 17 Visual
+
 ```{r echo=FALSE}
@@
 
-summary_table_18 %>% gt()
+summary_table_18 %>% 
+  gt() %>% 
+  fmt_percent(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Mean", "StDev", "Max", "Correlation"), 
+    decimals = 2
+  ) %>% 
+  fmt_number(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Sharpe", "TStat"), 
+    decimals = 3 
+  ) %>% 
+  fmt_number(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Obs"), 
+    decimals = 0 
+  )
 
@@
 # Slide 18 Visual
+
 ```{r echo=FALSE}
@@
 ```
+
 # Slide 21 Table (Im Confused)
+
 ```{r echo=FALSE}
-slide21_data <- port_returns_annualized %>% 
-  select(DATE,RET_LongPosV1,RET_LongPosV2,RET_ShortPosV1,RET_ShortPosV2) %>% 
-  mutate(
-    BaseLong = RET_LongPosV1,
-    BaseShort = RET_ShortPosV1,
-    NewLong = RET_LongPosV2,
-    NewShort = RET_ShortPosV2
+slide21_data_IVOL <- port_df_wide %>%
+  select(DATE, IVOL_LongPosV1, IVOL_LongPosV2, IVOL_ShortPosV1, IVOL_ShortPosV2) %>%
+  mutate(Year = year(DATE), Month = month(DATE)) %>%  
+  filter(Month == 12) %>%  
+  group_by(Year) %>% 
+  slice_tail(n = 1) %>%  #
+  ungroup()
+
+slide21_summary_table <- slide21_data_IVOL %>%
+  select(IVOL_LongPosV1,IVOL_LongPosV2, IVOL_ShortPosV1, IVOL_ShortPosV2) %>% 
+  rename_with(~ str_remove(.x, "IVOL_"), starts_with("IVOL_")) %>% 
+  summarise(
+    across(
+      c(LongPosV1, LongPosV2, ShortPosV1, ShortPosV2), 
+      list(
+        Mean = ~ mean(.x, na.rm = TRUE),
+        StDev = ~ sd(.x, na.rm = TRUE),
+        Max = ~ max(.x, na.rm = TRUE),
+        Min = ~ min(.x, na.rm = TRUE)
+      )
+    )
+  ) %>%
+  pivot_longer(
+    cols = everything(), 
+    names_to = c("Column", "Metric"), 
+    names_sep = "_", 
+    values_to = "Value"
+  ) %>%
+  pivot_wider(
+    names_from = Column, 
+    values_from = Value
   )
 
-slide21_table <- slide21_data %>% 
+slide21_summary_table %>%
+  gt() %>%
+  fmt_number(
+    columns = where(is.numeric), 
+    decimals = 2 
+  )
+```
+
+```{r echo=FALSE}
+slide21_data_SKEW <- port_df_wide %>%
+  select(DATE, SKEW_LongPosV1,SKEW_LongPosV2,SKEW_ShortPosV1,SKEW_ShortPosV2) %>%
+  mutate(Year = year(DATE), Month = month(DATE)) %>%  
+  filter(Month == 12) %>%  
+  group_by(Year) %>% 
+  slice_tail(n = 1) %>%  
+  ungroup()
+
+slide21_summary_table_SKEW <- slide21_data_SKEW %>% 
+  select(SKEW_LongPosV1,SKEW_LongPosV2,SKEW_ShortPosV1,SKEW_ShortPosV2) %>% 
+  rename_with(~ str_remove(.x, "SKEW_"), starts_with("SKEW_")) %>% 
   summarise(
-    across(c(BaseLong, BaseShort, NewLong, NewShort), list(
+    across(
+      c(LongPosV1, LongPosV2, ShortPosV1, ShortPosV2), 
+      list(
         Mean = ~ mean(.x, na.rm = TRUE),
         StDev = ~ sd(.x, na.rm = TRUE),
-      Min = ~ min(.x, na.rm = TRUE),
-      Max = ~ max(.x, na.rm = TRUE))))
+        Max = ~ max(.x, na.rm = TRUE),
+        Min = ~ min(.x, na.rm = TRUE)
+      )
+    )
+  ) %>%
+  pivot_longer(
+    cols = everything(), 
+    names_to = c("Column", "Metric"), 
+    names_sep = "_", 
+    values_to = "Value"
+  ) %>%
+  pivot_wider(
+    names_from = Column, 
+    values_from = Value
+  )
 
-summary_table_21 <- slide21_table %>%
+slide21_summary_table_SKEW %>%
+  gt() %>%
+  fmt_number(
+    columns = where(is.numeric), 
+    decimals = 2 
+  )
+```
+
+# Slide 23
+
+```{r echo=FALSE}
+
+symbol <- "SPY"
+start_date <- "2007-01-31"
+end_date <- "2022-12-31"
+
+spy_data <- tq_get(symbol, from = start_date, to = end_date)
+
+spy_monthly_returns <- spy_data %>%
+  tq_transmute(select = adjusted, 
+               mutate_fun = periodReturn, 
+               period = "monthly", 
+               type = "log") %>% 
+  mutate(SPY.RETURNS = monthly.returns)
+
+```
+
+```{r echo=FALSE}
+
+rf_symbol <- "DGS3MO"
+
+start_date <- "2007-01-31"
+end_date <- "2022-12-30"
+
+rf_data <- tq_get(rf_symbol, from = start_date, to = end_date, get = "economic.data")
+
+rf_monthly <- rf_data %>%
+  mutate(date = as.Date(date)) %>%
+  group_by(year = format(date, "%Y"), month = format(date, "%m")) %>%
+  slice_tail(n = 1) %>%  
+  ungroup() %>%
+  select(date, price) %>%
+  rename(rf.rate = price)
+
+rf_monthly <- rf_monthly %>%
+  mutate(rf.rate = (rf.rate / 100) / 12)
+
+```
+
+```{r}
+
+BenchIVOL_df <- read.csv("https://raw.githubusercontent.com/boykowealth/TailRiskPremium_PortfolioConcept/refs/heads/main/Portfolios_Formed_on_RESVAR.csv")
+
+```
+
+Merge
+
+```{r echo=FALSE}
+
+slide14_data <- slide14_data %>% mutate(date = as.Date(DATE))
+spy_monthly_returns <- spy_monthly_returns %>% mutate(date = as.Date(date))
+rf_monthly <- rf_monthly %>% mutate(date = as.Date(date))
+
+slide14_selected <- slide14_data %>% select(date, New, Base)
+spy_selected <- spy_monthly_returns %>% select(date, SPY.RETURNS)
+rf_selected <- rf_monthly %>% select(date, rf.rate)
+
+slide23_data <- slide14_selected %>%
+  left_join(spy_selected, by = "date") %>%
+  left_join(rf_selected, by = "date")
+
+slide23_data <- slide23_data %>% 
+  summarise(
+    across(c(New, Base,SPY.RETURNS, rf.rate ), list(
+      Mean = ~ mean(.x, na.rm = TRUE),
+      StDev = ~ sd(.x, na.rm = TRUE),
+      Max = ~ max(.x, na.rm = TRUE),
+      Obs = ~ sum(!is.na(.x)),
+      TStat = ~ mean(.x, na.rm = TRUE) / (sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x)))),
+      Sharpe = ~ mean(.x, na.rm = TRUE) / sd(.x, na.rm = TRUE)
+    )),
+    Correlation = cor(Base, New, use = "complete.obs")
+  )
+
+summary_table_23 <- slide23_data %>%
   pivot_longer(everything(), names_to = c("Column", "Metric"), names_sep = "_", values_to = "Value") %>%
@@
 
-summary_table_21 %>% gt()
+summary_table_23 %>% 
+   gt() %>% 
+  fmt_percent(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Mean", "StDev", "Max", "Correlation"), 
+    decimals = 2
+  ) %>% 
+  fmt_number(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Sharpe", "TStat"), 
+    decimals = 3 
+  ) %>% 
+  fmt_number(
+    columns = where(is.numeric), 
+    rows = Metric %in% c("Obs"), 
+    decimals = 0 
+  )
+
 ```

### 2025-03-20 416 Tables, Nearly done.qmd
Fixed Data

@@
 library(devtools)
+
+library(dplyr)
+library(tidyr)
 
@@
   mutate(SPY.RETURNS = monthly.returns)
-
 ```
@@
 
-```{r}
-
-BenchIVOL_df <- read.csv("https://raw.githubusercontent.com/boykowealth/TailRiskPremium_PortfolioConcept/refs/heads/main/Portfolios_Formed_on_RESVAR.csv")
-
+```{}
 ```
@@
 ```{r echo=FALSE}
+
+##
+BenchIVOL_df <- read.csv("FINAL_NOT_SUCKY.csv", colClasses = c("date" = "character")) %>% 
+  mutate(date = as.Date(zoo::as.yearmon(date, "%Y%m"), frac = 1),
+         hi20 = hi20 / 100,
+         lo20 = lo20 / 100,
+         fammaPort = (hi20 + lo20) / 2
+         ) %>% 
+  dplyr::select(date, fammaPort)
+
+##
 
@@
 slide23_data <- slide23_data %>% 
+  dplyr::left_join(BenchIVOL_df, by = "date")
+
+slide23_data <- slide23_data %>% 
   summarise(
-    across(c(New, Base,SPY.RETURNS, rf.rate ), list(
+    across(c(New, Base, fammaPort, SPY.RETURNS, rf.rate), list(
       Mean = ~ mean(.x, na.rm = TRUE),

### 2025-03-24 portfolio-monthly.py
Added Monthly Rebalancing

@@ -1,0 +1,210 @@
+import pandas as pd
+import numpy as np
+
+def data():
+    dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
+    dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
+    df = pd.read_csv(dataFile)
+
+    df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
+    df = df.sort_values(by='DATE', ascending=True)
+    df['MONTH'] = df['DATE'].dt.month
+    df['YEAR'] = df['DATE'].dt.year
+    df['YEAR_MONTH'] = df['DATE'].dt.to_period('M').dt.to_timestamp()
+
+    def convert_to_numeric(column):
+        return pd.to_numeric(column.str.replace('%', ''), errors='coerce')
+    
+    df['RET'] = convert_to_numeric(df['RET']) / 100
+    df['IVOL'] = convert_to_numeric(df['ivol']) / 100
+
+    df = df[['DATE','YEAR_MONTH','TICKER', 'RET', 'IVOL']]
+
+    return df
+
+def metaData(select=None):
+    df = data()
+    if select == 'tickerCount':
+        ticker_count = df['TICKER'].nunique()
+        return ticker_count
+    
+    elif select == 'tickers':
+        tickers = list(df['TICKER'].unique())
+        return tickers
+    
+    elif select == 'years':
+        years = list(df['YEAR_MONTH'].unique())
+        return years
+
+def portfolio():
+    df = data()
+    v1_list = [] ## Cols: Year, LongPosV1, ShortPosV1, LongRetV1, ShortRetV1, LongSkewV1, ShortSkewV1
+    
+    rebalance_period = metaData(select='years')
+
+    ### + Add Random Sample
+
+    ## V1->V2 Screen (Extreme IVOL High-Low) -> (Extreme SKEW Positive-Negative)
+    for period in rebalance_period:
+        period_timestamp = pd.to_datetime(period)
+        df_back = df[(df['YEAR_MONTH'] <= period_timestamp - pd.DateOffset(months=1)) & 
+             (df['YEAR_MONTH'] >= period_timestamp - pd.DateOffset(months=37))].copy()
[…]

### 2025-03-25 paper.qmd
Added Updates To Report

@@
 
+#### Monthly Rebalance Periods
+
+Majority of papers discussing the lottery preference premium, test their findings under the assumption of monthly rebalance periods.
+To test our strategy of annual rebalance, an additional test was completed to compare the growth of one dollar across each portfolio; under the assumption of monthly rebalance periods (all else equal).
+
+```{python}
+from sumStats import monthly_rebalance_compare
+
+monthly_rebalance_compare()
+
+```
+
 #### Explantion For V2 Short Measure Underperformance
 
+__Does high IVOL really describe a company's potential for valuation decline?__
+
+The paper "High Idiosyncratic Volatility and Low Returns" by Ang, Hodrick, Xing, and Zhang (2008) provides strong evidence of a negative relationship between past idiosyncratic volatility and future stock returns across 23 developed markets.
+
+__Summary of Key Findings:__
+
++ High IVOL → Low Returns: Stocks with high past idiosyncratic volatility earn significantly lower future returns, contradicting classic theories (like Merton 1987) that would expect higher risk to be rewarded.
++ Global Effect: This negative relationship is not U.S.-specific—it's observed in all G7 countries and persists after controlling for size, value, momentum, and market risk factors.
++ Robustness: The effect holds in equal- and value-weighted portfolios, with different volatility estimation windows, and remains unexplained by trading frictions, information asymmetry, skewness, or leverage interactions.
++ Magnitude: In the U.S., moving from the lowest to highest IVOL quintile is associated with a 0.73% drop in monthly returns.
+
+__Link to Valuation Decline:__
+
+While idiosyncratic volatility doesn’t cause valuation decline, the paper shows that high IVOL stocks tend to underperform, implying a greater probability of negative revaluation or weaker investor demand.
+So, IVOL can act as a proxy for valuation risk in practical terms.
+
+__"Idiosyncratic Volatility and Stock Returns: A Cross Country Analysis" by Visaltanachoti and Pukthuanthong (2009):__
+
+__Paper Summary:__ Using CAPM-based regressions across 38 countries (1996–2005), the authors examine whether idiosyncratic volatility (IVOL) is priced into stock returns.
+Only two countries (US and Philippines) show a positive relationship (higher IVOL → higher returns).
++ Ten countries (e.g., Australia, Canada, Japan, UK) show a negative relationship: higher IVOL is associated with lower stock returns.
++ The remaining countries show no significant relationship.
++ They conclude that IVOL is not consistently priced across global markets, making its effect context-dependent and potentially non-systematic.
+
+__IVOL & Valuation Decline__
+
+In countries where IVOL correlates negatively with returns, it implies that stocks with higher IVOL underperform, suggesting increased downside risk or investor aversion to uncertainty.
+This behavior aligns with the idea that higher IVOL may be interpreted as noise, mispricing, or overoptimism, leading to future valuation declines.
+So yes—in several markets, high IVOL is empirically linked to valuation decline, but not universally.
+
 #### Explanation For IVOL Stabalization Post 2008
+
+The number of companies covered by MSCI ESG Ratings doubled from 2014 to 2019. ESG fund assets rose over 250% from 2009 to 2019.
 
@@
 
-1. "Temporal Evolution of Financial Market Correlations" (Fenn et al.)
+__"Temporal Evolution of Financial Market Correlations" (Fenn et al.)__
 
@@
 
-2. "Correlation of Financial Markets in Times of Crisis" (Sandoval & Franca)
+__"Correlation of Financial Markets in Times of Crisis" (Sandoval & Franca)__
 
@@
 
-3. "Stock Markets Correlation - Before and During the Crisis Analysis" (Moldovan)​"
+__"Stock Markets Correlation - Before and During the Crisis Analysis" (Moldovan)​"__
 
@@
 The negative Dow-Nikkei correlation is an interesting insight, possibly reflecting a flight to safety in Japanese markets during the US downturn.
+
+__Bottom Line:__
+
+By 2012, consumer expectations for CSR became mainstream, driven by post-crisis distrust, social media activism, regulation, and millennial influence.
+This permanently reshaped corporate risk profiles, reinforcing CSR as a risk-mitigation tool in financial markets.
+
+__“The Impact of ESG Performance on Firm-Idiosyncratic Risk in the US and Canada” by Nutcha Kongpreecha (2021):__
+
++ Goal: Investigates whether ESG performance reduces firm-specific (idiosyncratic) risk across 480 listed U.S. and Canadian firms from 2007–2020.
++ Main Finding: Higher ESG scores are linked to lower idiosyncratic risk, especially during recessions and the COVID-19 pandemic.
+
+__Key Insights:__
+
++ ESG reduces idiosyncratic risk across both sensitive and non-sensitive industries, but environmental performance has an extra effect in sensitive sectors.
++ High market value firms benefit more from ESG than low-value ones.
++ Low-leverage firms gain more risk reduction from ESG than high-leverage ones.
++ The social pillar of ESG had the strongest effect during COVID-19, reflecting workforce protections and operational resilience.
+
+__Data & Methodology__
+
++ Sample: 480 publicly traded companies → 6,720 firm-year observations.
++ Data Source: Refinitiv Datastream (for ESG scores, stock data, and financials).
+
+__Idiosyncratic Risk Calculation:__
+
++ Uses Fama-French 3-Factor Model.
++ Daily residuals from regressions are annualized to compute idiosyncratic volatility (IDIOR)
+
+__Key Empirical Findings__
+
++ Main Result: A 1 standard deviation increase in ESG score (~20 points) reduces IDIOR by 1.56 percentage points, or ~9.8% of the standard deviation of IDIOR.
+This is statistically and economically significant.
++ No overall ESG score difference between sensitive vs.
+non-sensitive sectors.
+Environmental pillar has a stronger effect in sensitive industries (e.g., oil, chemicals) Social and Governance pillars did not show significant differences by sector.
+__Interpretation:__ Environmental risks (e.g., pollution penalties, regulations) are more pressing in these industries.
++ High Market Value firms benefit more from ESG in reducing IDIOR. Possibly due to greater stakeholder visibility, reputation effects, and more resources to implement ESG effectively.
++ Low-leverage firms see greater reduction in IDIOR from ESG. ESG may help mitigate asymmetric information for these firms lacking debt-market scrutiny.
++ ESG's risk-reduction effect is stronger in downturns (2008, 2009, 2020). During COVID-19, ESG reduced IDIOR by 4.48 percentage points, much more than in pre-COVID periods.
++ Only the social score had a significant added effect during COVID. __Interpretation:__ Stakeholder (employee/customer) treatment mattered most in a human-health crisis.
++ Stakeholder theory: ESG builds trust with employees, customers, and investors.
++ Legitimacy theory: Firms reduce reputational risk and regulatory pressure by conforming to social norms through ESG.
+
+Empirically, ESG is not just a “virtue signal” — it is quantitatively shown to lower firm-specific risk, especially in:
+
++ Environmentally exposed industries
++ Periods of market stress
++ Firms with fewer alternative risk buffers (e.g., low leverage)
+
+This supports our thesis that the post-2008 decline in idiosyncratic volatility may be linked to widespread ESG adoption, creating stronger correlation across firms and reducing return dispersion.

### 2025-03-25 portfolio.py

@@ -4,7 +4,7 @@ import numpy as np
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile2)
+    df = pd.read_csv(dataFile)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)

### 2025-03-25 portfolio_monthly.py

@@ -207,4 +207,4 @@ def monthly_total_return():
     plt.grid(True)
     plt.show()
 
-monthly_total_return()
+

### 2025-03-25 sumStats.py

@@ -1,6 +1,11 @@
 from portfolio import pchart
 
+from monthly_returns import portfolio_returns as yearly
+from portfolio_monthly import portfolio_returns as monthly
+
 import pandas as pd
+import matplotlib.pyplot as plt
+import numpy as np
 
 def tickers():
 
@@ -15,4 +20,36 @@ def tickers():
 
     return data
 
-tickers()
+def monthly_rebalance_compare():
+
+    dY = yearly()
+    dM = monthly()
+
+    dY = dY[['DATE', 'PORTFOLIO', 'RET']]
+    dM = dM[['DATE', 'PORTFOLIO', 'RET']]
+
+
+
+    d = pd.merge(dY, dM, how='left', on=['DATE', 'PORTFOLIO'])
+    d = d[['DATE', 'PORTFOLIO', 'RET_x', 'RET_y']]
+    d['CUM_RET_x'] = d.groupby(['PORTFOLIO'])['RET_x'].transform(lambda x: np.cumprod(1 + x))
+    d['CUM_RET_y'] = d.groupby(['PORTFOLIO'])['RET_y'].transform(lambda x: np.cumprod(1 + x))
+
+    portfolios = d['PORTFOLIO'].unique()
+
+    fig, axes = plt.subplots(2, 2, figsize=(8, 8))
+    axes = axes.flatten()
+
+    for i, portfolio in enumerate(portfolios):
+        ax = axes[i]
+        portfolio_data = d[d['PORTFOLIO'] == portfolio]
+        
+        ax.plot(portfolio_data['DATE'], portfolio_data['CUM_RET_x'], label='Annual Rebalance')
+        ax.plot(portfolio_data['DATE'], portfolio_data['CUM_RET_y'], label='Monthly Rebalance')
+        
+        ax.set_title(f'Portfolio Comparison {portfolio}')
+        ax.legend()
+        ax.grid(True)
+
+    plt.tight_layout()
+    plt.show()

### 2025-03-25 visuals.py

@@ -18,10 +18,12 @@ def monthly_ret():
         data = portfolio[portfolio['PORTFOLIO'] == port]
         ax = axes[i]
         ax.plot(data['DATE'], data['RET'], label=port)
-        ax.set_xlabel("Date")
-        ax.set_ylabel("Return")
         ax.set_title(f"Portfolio: {port}")
         ax.yaxis.set_major_formatter(mtick.PercentFormatter(1.0))
+        ax.set_xlabel("Date", fontsize=8)
+        ax.set_ylabel("Return", fontsize=8)
+        ax.tick_params(axis='x', labelsize=8)
+        ax.tick_params(axis='y', labelsize=8)
         ax.legend()
         ax.grid(True)
 

### 2025-03-25 416 Tables Complete.qmd
Add files via upload

--- 2025-03-20 416 Tables, Nearly done.qmd
@@
+![](images/clipboard-1227717430.png)
+
 ---
@@
       Max = ~ max(.x, na.rm = TRUE),
+      Min = ~ min(.x, na.rm = TRUE),
       Obs = ~ sum(!is.na(.x)),
@@
    gt() %>% 
+  cols_label(
+    NewSubBase = "New - Base"
+  ) %>%
   fmt_percent(
@@
   gt() %>% 
+  cols_label(
+    BaseLong = "Base Long",
+    NewLong = "New Long",
+    NewSubBaseLong = "New Long -  Base Long"
+
+  ) %>%
   fmt_percent(
@@
   gt() %>%
+  cols_label(
+
+    BaseShort = "Base Short",
+    NewShort = "New Short",
+    NewSubBaseShort = "New - Base"
+  ) %>%
+
   fmt_percent(
@@
 slide21_summary_table %>%
-  gt() %>%
+  gt() %>%cols_label(
+    LongPosV1 = "Base IVOL Long",
+    LongPosV2 = "New IVOL Long",
+    ShortPosV1 = "Base IVOL Short",
+    ShortPosV2 = "New IVOL Short"
+  ) %>%
   fmt_number(
@@
   gt() %>%
+  cols_label(
+    LongPosV1 = "Base Skew Long",
+    LongPosV2 = "New Skew Long",
+    ShortPosV1 = "Base Skew Short",
+    ShortPosV2 = "New  Skew Short"
+  ) %>%
+
   fmt_number(
@@
 
-```{}
+```         
 ```
@@
 ##
-BenchIVOL_df <- read.csv("FINAL_NOT_SUCKY.csv", colClasses = c("date" = "character")) %>% 
+BenchIVOL_df <- read.csv("https://raw.githubusercontent.com/boykowealth/TailRiskPremium_PortfolioConcept/refs/heads/main/FINAL_NOT_SUCKY.csv", colClasses = c("date" = "character")) %>% 
   mutate(date = as.Date(zoo::as.yearmon(date, "%Y%m"), frac = 1),
@@
 
-slide23_data <- slide23_data %>% 
+# Compute Summary Stats
+slide23_summary <- slide23_data %>% 
   summarise(
@@
       Sharpe = ~ mean(.x, na.rm = TRUE) / sd(.x, na.rm = TRUE)
-    )),
-    Correlation = cor(Base, New, use = "complete.obs")
+    ))
   )
 
-summary_table_23 <- slide23_data %>%
+# Add Correlation rows manually
+correlations <- slide23_data %>%
+  summarise(
+    Base = cor(New, Base, use = "complete.obs"),
+    fammaPort = cor(New, fammaPort, use = "complete.obs"),
+    SPY.RETURNS = cor(New, SPY.RETURNS, use = "complete.obs"),
+    rf.rate = cor(New, rf.rate, use = "complete.obs"),
+    New = 1  # Perfect correlation with self
+  ) %>%
+  pivot_longer(everything(), names_to = "Column", values_to = "Value") %>%
+  mutate(Metric = "Correlation")
+
+# Calculate difference stats vs. New
+diff_stats <- slide23_data %>%
+  summarise(
+    across(c(Base, fammaPort, SPY.RETURNS, rf.rate), 
+           list(
+             DiffMean = ~ mean(New - .x, na.rm = TRUE),
+             DiffTStat = ~ {
+               diff <- New - .x
+               mean(diff, na.rm = TRUE) / (sd(diff, na.rm = TRUE) / sqrt(sum(!is.na(diff))))
+             }
+           ), .names = "{.col}_{.fn}")
+  ) %>%
+  pivot_longer(everything(), 
+               names_to = c("Column", "Metric"), 
+               names_sep = "_", 
+               values_to = "Value") %>%
+  mutate(Metric = case_when(
+    Metric == "DiffMean" ~ "Mean of Difference w.r.t. New",
+    Metric == "DiffTStat" ~ "T-Stat of Difference w.r.t. New"
+  )) %>%
+  pivot_wider(names_from = Column, values_from = Value)
+
+# Tidy + Pivot for GT table
+summary_table_23 <- slide23_summary %>%
   pivot_longer(everything(), names_to = c("Column", "Metric"), names_sep = "_", values_to = "Value") %>%
+  bind_rows(correlations) %>% 
   pivot_wider(names_from = Column, values_from = Value)
+summary_table_23 <- bind_rows(summary_table_23, diff_stats)
 
@@
   gt() %>% 
+  cols_label(
+    New = "New Strategy",
+    Base = "Base Strategy",
+    fammaPort = "Fama-French Benchmark",
+    SPY.RETURNS = "S&P 500 Returns",
+    rf.rate = "Risk-Free Rate"
+  ) %>% 
   fmt_percent(
     columns = where(is.numeric), 
-    rows = Metric %in% c("Mean", "StDev", "Max", "Correlation"), 
+    rows = Metric %in% c("Mean", "StDev", "Max", "Correlation", "Mean of Difference w.r.t. New"), 
     decimals = 2
@@
     columns = where(is.numeric), 
-    rows = Metric %in% c("Sharpe", "TStat"), 
+    rows = Metric %in% c("Sharpe", "TStat", "T-Stat of Difference w.r.t. New"), 
     decimals = 3 

### 2025-03-25 paper.qmd
Updated Report To have Tables

--- 2025-03-18 paper.qmd
@@
 ---
+
 ## Abstract
+
 The investment strategy centered around the Idiosyncratic Volatility (IVOL) Risk Premium exploits the distinct performance patterns of stocks exhibiting varying levels of firm-specific risk.
@@
 ## Investment Strategy
+
 The core of the strategy involves taking long positions in stocks characterized by low idiosyncratic volatility, which consistently exhibit a positive return premium.
@@
 ## Persistence of the Idiosyncratic Risk Premium:
+
 The persistence of the idiosyncratic risk premium is attributed to several factors, including investor behavior, market frictions, and liquidity constraints.
@@
 ## Strategy Details
+
 To effectively harness the Idiosyncratic Volatility (IVOL) Risk Premium, the strategy employs a systematic approach in selecting and managing investments:
@@
 ### Data Selection:
-+ The initial step involves gathering comprehensive data from the WRDS database by Wharton, spanning from 2002 to 2022.
+
+-   The initial step involves gathering comprehensive data from the WRDS database by Wharton, spanning from 2002 to 2022.
 This extensive data set includes stock prices, historical returns, and idiosyncratic volatility measures.
-+ The portfolio will be rebalanced annually from 2007 to 2022, ensuring a lookback window of five additional years to refine the selection criteria.
+-   The portfolio will be rebalanced annually from 2007 to 2022, ensuring a lookback window of five additional years to refine the selection criteria.
 
 ### Portfolio Construction
-+ Selection Criteria: The strategy identifies stocks based on idiosyncratic volatility and skewness.
+
+-   Selection Criteria: The strategy identifies stocks based on idiosyncratic volatility and skewness.
 Stocks exhibiting high idiosyncratic volatility and negative skewness are shortlisted for short positions, while those with low idiosyncratic volatility and positive skewness are chosen for long positions.
 
-+ Long Positions: The strategy takes long positions in stocks with low idiosyncratic volatility, aiming to capitalize on their consistent positive return premium.
+-   Long Positions: The strategy takes long positions in stocks with low idiosyncratic volatility, aiming to capitalize on their consistent positive return premium.
 These stocks are often underpriced relative to their risk profiles, offering reliable opportunities for excess returns.
 
-+ Short Positions: Simultaneously, short positions are established in stocks with high idiosyncratic volatility.
+-   Short Positions: Simultaneously, short positions are established in stocks with high idiosyncratic volatility.
 These stocks are systematically mispriced due to behavioral biases and market frictions, and they tend to underperform over time.
@@
 ### Risk Management:
-+ Diversification: The portfolio is diversified across multiple stocks to mitigate firm-specific risks.
+
+-   Diversification: The portfolio is diversified across multiple stocks to mitigate firm-specific risks.
 By constructing a well-balanced portfolio, the strategy ensures that the impact of any single stock's volatility is minimized.
-+ Rebalancing: Annual rebalancing is conducted to maintain the desired risk exposure and adjust for changes in idiosyncratic volatility and skewness.
+-   Rebalancing: Annual rebalancing is conducted to maintain the desired risk exposure and adjust for changes in idiosyncratic volatility and skewness.
 This regular review allows the strategy to adapt to evolving market conditions and optimize returns.
@@
 ### Performance Monitoring:
-+ The strategy's performance is continuously monitored against relevant benchmarks to ensure alignment with its objectives.
+
+-   The strategy's performance is continuously monitored against relevant benchmarks to ensure alignment with its objectives.
 Key performance indicators, such as risk-adjusted returns and volatility measures, are tracked to assess the strategy's effectiveness.
-+ Adjustments: Based on performance analysis, adjustments may be made to the portfolio composition, selection criteria, or rebalancing frequency to enhance the strategy's overall performance and resilience to market fluctuations.
+-   Adjustments: Based on performance analysis, adjustments may be made to the portfolio composition, selection criteria, or rebalancing frequency to enhance the strategy's overall performance and resilience to market fluctuations.
 
 ## Performance Evaluations
+
 A variety of benchmark statistics will be included in our portfolio testing:
 
-+ Annual / Monthly Return Values/Charts For 4 Portfolios
-+ Description of Stock Breakdown For Four Portfolios (Skew, IVOL)
-+ V1 & V2 Screening
+-   Annual / Monthly Return Values/Charts For 4 Portfolios
+-   Description of Stock Breakdown For Four Portfolios (Skew, IVOL)
+-   V1 & V2 Screening
 
@@
 V2 portfolios describe our teams extension of the V1 sceen, which utilizes skew to screen for 30 companies in each portfolio.
+
+#### New vs. Base Overall Strategy Return Summary Statistics & Hypothesis Test
+
+```{python}
+from sumStats import htmlTable
+
+htmlTable("slide14_summary.csv")
+```
 
@@
 
-#### Cumalitve Monthly Returns (Value of $1 Invested)
+#### Cumalitve Monthly Returns (Value of \$1 Invested)
 
@@
 
-#### Annual Returns (Value of $1 Invested)
+#### New vs. Base Long Side Return Summary Statistics & Hypothesis Test
+
+```{python}
+from sumStats import htmlTable
+
+htmlTable("slide17_summary.csv")
+```
+
+#### New vs. Base Short Side Return Summary Statistics & Hypothesis Test
+
+```{python}
+from sumStats import htmlTable
+
+htmlTable("slide18_summary.csv")
+```
+
+#### Annual Returns (Value of \$1 Invested)
+
 ```{python}
@@
 In contrast, the Base Short utilizes the highest 100 IVOL from the dataset.
+
+#### Portfolio Attributes: Summary Statistics of Screening Variables (IVOL)
+
+```{python}
+from sumStats import htmlTable
+
+htmlTable("slide21_ivol_summary.csv")
+```
+
+#### Portfolio Attributes: Summary Statistics of Screening Variables (Skew)
+
+```{python}
+from sumStats import htmlTable
+
+htmlTable("slide21_skew_summary.csv")
+```
 
@@
 ### Portfolio Positions
+
 The portfolios are equal weighted. This provides an accurate representation of both the V1 and V2 screening effectiveness.
+
+#### Benchmark Analysis: Return Summary Statistics
+
+```{python}
+from sumStats import htmlTable
+
+htmlTable("slide23_comparison_summary.csv")
+```
 
@@
 
-#### Equal Weighted Portfolios (Baseline vs Strategy-Extension) Cumalitve Returns (Value of $1 Invested)
+#### Equal Weighted Portfolios (Baseline vs Strategy-Extension) Cumalitve Returns (Value of \$1 Invested)
 
@@
 ```
+
+### Additional Tests
+
+A series of additional tests were completed to better explain the results found in our data sample period. These tests were conducted in order to improve upon the strategy.
+
+#### Monthly Rebalance Periods
+
+Majority of papers discussing the lottery preference premium, test their findings under the assumption of monthly rebalance periods.
+To test our strategy of annual rebalance, an additional test was completed to compare the growth of one dollar across each portfolio; under the assumption of monthly rebalance periods (all else equal).
+
+```{python}
+from sumStats import monthly_rebalance_compare
+
+monthly_rebalance_compare()
+
+```
+
+#### Explantion For V2 Short Measure Underperformance
+
+**Does high IVOL really describe a company's potential for valuation decline?**
+
+The paper "High Idiosyncratic Volatility and Low Returns" by Ang, Hodrick, Xing, and Zhang (2008) provides strong evidence of a negative relationship between past idiosyncratic volatility and future stock returns across 23 developed markets.
+
+**Summary of Key Findings:**
+
+-   High IVOL → Low Returns: Stocks with high past idiosyncratic volatility earn significantly lower future returns, contradicting classic theories (like Merton 1987) that would expect higher risk to be rewarded.
+-   Global Effect: This negative relationship is not U.S.-specific—it's observed in all G7 countries and persists after controlling for size, value, momentum, and market risk factors.
+-   Robustness: The effect holds in equal- and value-weighted portfolios, with different volatility estimation windows, and remains unexplained by trading frictions, information asymmetry, skewness, or leverage interactions.
+-   Magnitude: In the U.S., moving from the lowest to highest IVOL quintile is associated with a 0.73% drop in monthly returns.
+
+**Link to Valuation Decline:**
+
+While idiosyncratic volatility doesn’t cause valuation decline, the paper shows that high IVOL stocks tend to underperform, implying a greater probability of negative revaluation or weaker investor demand.
+So, IVOL can act as a proxy for valuation risk in practical terms.
+
+**"Idiosyncratic Volatility and Stock Returns: A Cross Country Analysis" by Visaltanachoti and Pukthuanthong (2009):**
+
+**Paper Summary:** Using CAPM-based regressions across 38 countries (1996–2005), the authors examine whether idiosyncratic volatility (IVOL) is priced into stock returns.
+Only two countries (US and Philippines) show a positive relationship (higher IVOL → higher returns).
++ Ten countries (e.g., Australia, Canada, Japan, UK) show a negative relationship: higher IVOL is associated with lower stock returns.
++ The remaining countries show no significant relationship.
++ They conclude that IVOL is not consistently priced across global markets, making its effect context-dependent and potentially non-systematic.
+
+**IVOL & Valuation Decline**
+
+In countries where IVOL correlates negatively with returns, it implies that stocks with higher IVOL underperform, suggesting increased downside risk or investor aversion to uncertainty.
+This behavior aligns with the idea that higher IVOL may be interpreted as noise, mispricing, or overoptimism, leading to future valuation declines.
+So yes—in several markets, high IVOL is empirically linked to valuation decline, but not universally.
+
+#### Explanation For IVOL Stabalization Post 2008
+
+The number of companies covered by MSCI ESG Ratings doubled from 2014 to 2019. ESG fund assets rose over 250% from 2009 to 2019.
+
+**Key Takeaways for Post-2008 Market Stability**
+
+-   Systemic risk increased post-2008, making diversification ineffective.
+-   Correlations tend to rise before a crisis, suggesting a leading indicator for instability.
+-   Investor behavior shifted towards liquidity and safety, aligning with our discussion on idiosyncratic risk vs. systemic risk.
+-   Market-wide herding intensified, driving the preference for safe assets and reducing the appeal of high-idiosyncratic-risk stocks.
+-   Relative decline in idiosyncratic risk due to rising market and industry volatility.
+-   Crisis events (2008, COVID-19) still caused firm-specific volatility spikes but within an environment of elevated systematic risk.
+-   Investor preference for stability (CSP, large firms) helped reduce firm-specific risk, supporting our discussion on flight to safety and liquidity-driven volatility shifts.
+
+**"Temporal Evolution of Financial Market Correlations" (Fenn et al.)**
+
+**Summary:** This study applies Principal Component Analysis (PCA) and Random Matrix Theory (RMT) to analyze evolving correlations in financial markets from 1999–2010.
+Key findings indicate that post-2008, financial assets across different markets exhibited much stronger correlations.
+The first principal component accounted for nearly 40% of market variance post-Lehman collapse, suggesting that most assets were moving together.
+This heightened correlation reduces diversification benefits and increases systemic risk.
+
+**Relevance to Post-2008 Stability:** This aligns with our discussion on correlation risk, where increased asset co-movement reduces the effectiveness of diversification.
+It also supports the high systemic risk in a post-2008 world, showing why investor behavior shifted towards liquidity preference and safe-haven assets.
+
+**"Correlation of Financial Markets in Times of Crisis" (Sandoval & Franca)**
+
+**Summary:** The study uses eigenvalue decomposition of correlation matrices to show that market correlations increase dramatically during crises (1987, 1998, 2001, 2008).
+Their key finding is that high volatility leads to strong correlations, meaning that during crashes, markets behave as one.
+The study also finds that average market correlation increases prior to a crisis, suggesting an early warning signal.
+
+**Relevance to Post-2008 Stability:** This reinforces our point that in crisis conditions, herding behavior dominates.
+The finding that correlations increase before crises aligns with the liquidity crunch hypothesis, where investors pull out of riskier assets preemptively.
+The study also supports the behavioral premium idea, where investors chase "lottery stocks" in stable times but rush to liquidity in crises.
+
+**"Stock Markets Correlation - Before and During the Crisis Analysis" (Moldovan)​"**
+
+**Summary:** This paper specifically analyzes the correlations between New York (Dow Jones), London (FTSE), and Tokyo (Nikkei) before (2003-2006) and during (2007-2010) the crisis.
+It finds that correlations were significantly stronger during the crisis period, meaning that stocks moved together more in downturns than in uptrends.
+Regression models showed that FTSE’s influence on Dow and Nikkei increased during the crisis, while Nikkei and Dow became negatively correlated.
+
+**Relevance to Post-2008 Stability:** This supports our discussion on global financial interdependence post-2008.
+The increased correlation in downturns confirms that diversification failed when most needed.
+The negative Dow-Nikkei correlation is an interesting insight, possibly reflecting a flight to safety in Japanese markets during the US downturn.
+
+**Bottom Line:**
+
+By 2012, consumer expectations for CSR became mainstream, driven by post-crisis distrust, social media activism, regulation, and millennial influence.
+This permanently reshaped corporate risk profiles, reinforcing CSR as a risk-mitigation tool in financial markets.
+
+**“The Impact of ESG Performance on Firm-Idiosyncratic Risk in the US and Canada” by Nutcha Kongpreecha (2021):**
+
+-   Goal: Investigates whether ESG performance reduces firm-specific (idiosyncratic) risk across 480 listed U.S. and Canadian firms from 2007–2020.
+-   Main Finding: Higher ESG scores are linked to lower idiosyncratic risk, especially during recessions and the COVID-19 pandemic.
+
+**Key Insights:**
+
+-   ESG reduces idiosyncratic risk across both sensitive and non-sensitive industries, but environmental performance has an extra effect in sensitive sectors.
+-   High market value firms benefit more from ESG than low-value ones.
+-   Low-leverage firms gain more risk reduction from ESG than high-leverage ones.
+-   The social pillar of ESG had the strongest effect during COVID-19, reflecting workforce protections and operational resilience.
+
+**Data & Methodology**
+
+-   Sample: 480 publicly traded companies → 6,720 firm-year observations.
+-   Data Source: Refinitiv Datastream (for ESG scores, stock data, and financials).
+
+**Idiosyncratic Risk Calculation:**
+
+-   Uses Fama-French 3-Factor Model.
+-   Daily residuals from regressions are annualized to compute idiosyncratic volatility (IDIOR)
+
+$$
+R_{i, d} - R_{f, d} = \alpha_i + \beta_i (R_{m, d} - R_{f, d}) + s_i SMB_d + h_i HML_d + u_{i, d}
+$$
+
+**Key Empirical Findings**
+
+-   Main Result: A 1 standard deviation increase in ESG score (\~20 points) reduces IDIOR by 1.56 percentage points, or \~9.8% of the standard deviation of IDIOR.
+This is statistically and economically significant.
+-   No overall ESG score difference between sensitive vs.
+non-sensitive sectors.
+Environmental pillar has a stronger effect in sensitive industries (e.g., oil, chemicals) Social and Governance pillars did not show significant differences by sector.
+**Interpretation:** Environmental risks (e.g., pollution penalties, regulations) are more pressing in these industries.
+-   High Market Value firms benefit more from ESG in reducing IDIOR. Possibly due to greater stakeholder visibility, reputation effects, and more resources to implement ESG effectively.
+-   Low-leverage firms see greater reduction in IDIOR from ESG. ESG may help mitigate asymmetric information for these firms lacking debt-market scrutiny.
+-   ESG's risk-reduction effect is stronger in downturns (2008, 2009, 2020). During COVID-19, ESG reduced IDIOR by 4.48 percentage points, much more than in pre-COVID periods.
+-   Only the social score had a significant added effect during COVID. **Interpretation:** Stakeholder (employee/customer) treatment mattered most in a human-health crisis.
+-   Stakeholder theory: ESG builds trust with employees, customers, and investors.
+-   Legitimacy theory: Firms reduce reputational risk and regulatory pressure by conforming to social norms through ESG.
+
+Empirically, ESG is not just a “virtue signal” — it is quantitatively shown to lower firm-specific risk, especially in:
+
+-   Environmentally exposed industries
+-   Periods of market stress
+-   Firms with fewer alternative risk buffers (e.g., low leverage)
+
+This supports our thesis that the post-2008 decline in idiosyncratic volatility may be linked to widespread ESG adoption, creating stronger correlation across firms and reducing return dispersion.

### 2025-03-25 sumStats.py

@@ -53,3 +53,10 @@ def monthly_rebalance_compare():
 
     plt.tight_layout()
     plt.show()
+
+def htmlTable(path):
+
+    df = pd.read_csv(path)
+    df = df.to_html(index=False, float_format="{:.4f}".format)
+
+    return df

### 2025-03-25 visuals.py

@@ -19,7 +19,7 @@ def monthly_ret():
         ax = axes[i]
         ax.plot(data['DATE'], data['RET'], label=port)
         ax.set_title(f"Portfolio: {port}")
-        ax.yaxis.set_major_formatter(mtick.PercentFormatter(1.0))
+        ## ax.yaxis.set_major_formatter(mtick.PercentFormatter(1.0))
         ax.set_xlabel("Date", fontsize=8)
         ax.set_ylabel("Return", fontsize=8)
         ax.tick_params(axis='x', labelsize=8)

### 2025-03-27 esg_scr.py
Added images

@@ -1,0 +1,138 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import matplotlib.dates as mdates
+import numpy as np
+from pandas.tseries.offsets import MonthBegin
+
+from monthly_returns import portfolio_returns
+
+esg = pd.read_csv(r"C:\Users\brayd\Downloads\multiTimeline (3).csv")
+csr = pd.read_csv(r"C:\Users\brayd\Downloads\multiTimeline (4).csv")
+ivol = portfolio_returns()
+ivol = ivol[ivol['PORTFOLIO'].isin(['LongPosV1', 'ShortPosV1'])]
+ivol = ivol.pivot(index='DATE', columns='PORTFOLIO', values='IVOL').reset_index()
+ivol['Value'] = (ivol['LongPosV1'] + ivol['ShortPosV1']) / 2
+ivol['DATE'] = pd.to_datetime(ivol['DATE']) - MonthBegin(1)
+ivol = ivol[['DATE', 'Value']]
+
+
+def ESG():
+    df = esg
+    df.columns = ['Month', 'Value']
+
+    df['Month'] = pd.to_datetime(df['Month'])
+    df = df.sort_values(by=['Month'])
+
+    fig, ax = plt.subplots(figsize=(8, 5))
+    fig.patch.set_facecolor('#1D1D1B')
+    ax.set_facecolor('#1D1D1B')
+
+    ax.plot(df['Month'], df['Value'], label='ESG Search', color='#9fc3cf')
+
+    ax.tick_params(axis='x', rotation=45) 
+
+    ax.spines['bottom'].set_color('white')
+    ax.spines['top'].set_color('white')
+    ax.spines['left'].set_color('white')
+    ax.spines['right'].set_color('white')
+
+    ax.tick_params(axis='x', colors='white')
+    ax.tick_params(axis='y', colors='white')
+
+    ax.set_xlabel("Month", fontsize=10, color='white')
+    ax.set_ylabel("Indexed Search", fontsize=10, color='white')
+    ax.set_title("Search Indexed Frequency For Term: ESG In The United States (Google Trends)",
+                 fontsize=12, color='white', loc='left')
+    ax.grid(True, linestyle=':', color='white')
+
+    plt.tight_layout()
+    plt.show()
+
+def CSR():
+    df = csr
+    df.columns = ['Month', 'Value']
+
+    df['Month'] = pd.to_datetime(df['Month'])
+    df = df.sort_values(by=['Month'])
+
[…]

### 2025-03-27 portfolio.py

@@ -4,6 +4,7 @@ import numpy as np
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
+    dataFile3 = r"C:\Users\brayd\Downloads\kem6kumzrbylvuz3.csv" ## Earlier Period For Testing
     df = pd.read_csv(dataFile2)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
@@ -87,7 +88,7 @@ def portfolio():
     v1_df = pd.concat(v1_list, ignore_index=True)
     v1_df = v1_df.pivot(index=['Year', 'TICKER'], columns='Position', values=['T_RETURN', 'SKEW']).reset_index()
     v1_df.columns = ['Year', 'TICKER', 'LongRetV1', 'ShortRetV1', 'LongSkewV1', 'ShortSkewV1']
-    v1_df = v1_df[v1_df['Year'] >= 2007] ## Rebalances Begin In 2007
+    v1_df = v1_df[v1_df['Year'] >= 1965] ## Rebalances Begin In 2007
     rebalance_years = list(v1_df['Year'].unique())
 
     v2_list = []
@@ -142,3 +143,4 @@ def portfolio():
 
 def pchart():
     return portfolio()
+

### 2025-03-27 portfolio_monthly.py

@@ -4,7 +4,8 @@ import numpy as np
 def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
-    df = pd.read_csv(dataFile2)
+    dataFile3 = r"C:\Users\brayd\Downloads\kem6kumzrbylvuz3.csv"
+    df = pd.read_csv(dataFile3)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)
@@ -35,6 +36,7 @@ def metaData(select=None):
     elif select == 'years':
         years = list(df['YEAR_MONTH'].unique())
         return years
+
 
 def portfolio():
     df = data()

### 2025-03-27 visuals.py

@@ -13,20 +13,25 @@ def monthly_ret():
     unique_ports = portfolio['PORTFOLIO'].unique()
     
     fig, axes = plt.subplots(2, 2, figsize=(8, 8))
+    fig.patch.set_facecolor('#1D1D1B')
     axes = axes.flatten() 
     for i, port in enumerate(unique_ports[:4]):  
         data = portfolio[portfolio['PORTFOLIO'] == port]
         ax = axes[i]
-        ax.plot(data['DATE'], data['RET'], label=port)
-        ax.set_title(f"Portfolio: {port}")
-        ## ax.yaxis.set_major_formatter(mtick.PercentFormatter(1.0))
-        ax.set_xlabel("Date", fontsize=8)
-        ax.set_ylabel("Return", fontsize=8)
-        ax.tick_params(axis='x', labelsize=8)
-        ax.tick_params(axis='y', labelsize=8)
-        ax.legend()
-        ax.grid(True)
+        ax.plot(data['DATE'], data['RET'], label=port, color='#9fc3cf')
+        ax.set_title(f"Portfolio: {port}", color='white', fontsize=10, loc='left')
+        ax.set_xlabel("Date", fontsize=8, color='white')
+        ax.set_ylabel("Return", fontsize=8, color='white')
+        ax.tick_params(axis='x', labelsize=8, colors='white')
+        ax.tick_params(axis='y', labelsize=8, colors='white')
+        ax.grid(True, linestyle=':', color='white')
 
+        ax.yaxis.set_major_formatter(mtick.PercentFormatter(1.0))
+        ax.set_facecolor('#1D1D1B') 
+        ax.spines['top'].set_color('white')
+        ax.spines['right'].set_color('white')
+        ax.spines['bottom'].set_color('white')
+        ax.spines['left'].set_color('white')
 
     plt.tight_layout()  # Adjust spacing between plots
     plt.show()
@@ -36,6 +41,7 @@ def monthly_ivol():
     portfolio = portfolio[portfolio['PORTFOLIO'].isin(['LongPosV1', 'ShortPosV1'])]
     
     fig, axs = plt.subplots(2, 1, figsize=(8, 8))
+    fig.patch.set_facecolor('#1D1D1B')
     
     for i, port in enumerate(portfolio['PORTFOLIO'].unique()):
         data = portfolio[portfolio['PORTFOLIO'] == port]
@@ -51,17 +57,40 @@ def monthly_total_return():
     portfolio = portfolio_returns()
[…]

### 2025-03-29 models.py
Added Models

@@ -1,0 +1,32 @@
+import pandas as pd
+import numpy as np
+from scipy.stats import linregress
+
+from monthly_returns import portfolio_returns
+
+def ivol_ret():
+
+    ivol = portfolio_returns()
+
+    ivol = ivol[ivol['PORTFOLIO'].isin(['LongPosV1', 'ShortPosV1'])]
+
+    ivol_v = ivol.pivot(index='DATE', columns='PORTFOLIO', values='IVOL').reset_index()
+    print(ivol_v)
+    ivol_v['IVOL'] = (ivol_v['LongPosV1'] + ivol_v['ShortPosV1']) / 2
+    ivol_v['IVOL'] = ivol_v['IVOL'] - ivol_v['IVOL'].shift(1)
+    ivol_v = ivol_v[['DATE', 'IVOL']]
+
+    ivol_r = ivol.pivot(index='DATE', columns='PORTFOLIO', values='RET').reset_index()
+    ivol_r['RET'] = (ivol_r['LongPosV1'] + ivol_r['ShortPosV1']) / 2
+    ivol_r = ivol_r[['DATE', 'RET']]
+
+    df = pd.merge(ivol_v, ivol_r, on='DATE', how='left')
+    df = df[df['DATE'] >= "2007-01-01"]
+
+    results = linregress(df['IVOL'],df['RET'])
+
+    cor = np.corrcoef(df['IVOL'],df['RET'])
+
+    print(results, cor)
+
+ivol_ret()

### 2025-03-30 models.py
Added Charts For PowerPoint

@@ -1,32 +1,113 @@
 import pandas as pd
+from pandas.tseries.offsets import MonthEnd
+
 import numpy as np
+from numpy.polynomial.polynomial import Polynomial
 from scipy.stats import linregress
+
+import matplotlib.pyplot as plt
+import matplotlib.ticker as mticker
+from matplotlib.ticker import PercentFormatter
 
 from monthly_returns import portfolio_returns
 
+def benchmark():
+    df = pd.read_csv(r"C:\Users\Brayden Boyko\Downloads\HistoricalPrices (2).csv")
+
+    df['Date'] = pd.to_datetime(df['Date'])
+    df['MONTHYEAR'] = df['Date'].dt.to_period('M')
+    df['AVE'] = df.groupby('MONTHYEAR')[' Close'].transform('mean')
+    df = df.sort_values(by='Date', ascending=True)
+    df['DATE'] = df['Date']
+
+    df = df[['DATE', 'AVE']]
+
+    return df
+
 def ivol_ret():
-
     ivol = portfolio_returns()
+    bench = benchmark()
 
-    ivol = ivol[ivol['PORTFOLIO'].isin(['LongPosV1', 'ShortPosV1'])]
+    ivol = ivol[ivol['PORTFOLIO'].isin(['LongPosV1'])]
 
     ivol_v = ivol.pivot(index='DATE', columns='PORTFOLIO', values='IVOL').reset_index()
-    print(ivol_v)
-    ivol_v['IVOL'] = (ivol_v['LongPosV1'] + ivol_v['ShortPosV1']) / 2
-    ivol_v['IVOL'] = ivol_v['IVOL'] - ivol_v['IVOL'].shift(1)
+    ivol_v['IVOL'] = ivol_v['LongPosV1']
+    ivol_v['IVOL'] = np.log(ivol_v['IVOL'] / ivol_v['IVOL'].shift(1)) * 10
     ivol_v = ivol_v[['DATE', 'IVOL']]
 
     ivol_r = ivol.pivot(index='DATE', columns='PORTFOLIO', values='RET').reset_index()
-    ivol_r['RET'] = (ivol_r['LongPosV1'] + ivol_r['ShortPosV1']) / 2
+    ivol_r['RET'] = ivol_r['LongPosV1']
     ivol_r = ivol_r[['DATE', 'RET']]
 
     df = pd.merge(ivol_v, ivol_r, on='DATE', how='left')
[…]

### 2025-03-30 portfolio.py

@@ -5,7 +5,8 @@ def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile3 = r"C:\Users\brayd\Downloads\kem6kumzrbylvuz3.csv" ## Earlier Period For Testing
-    df = pd.read_csv(dataFile2)
+    dataFile4 = r"C:\Users\Brayden Boyko\Downloads\kem6kumzrbylvuz3.csv"
+    df = pd.read_csv(dataFile4)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)

### 2025-03-30 portfolio_monthly.py

@@ -5,7 +5,7 @@ def data():
     dataFile = r"C:\Users\Brayden Boyko\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile2 = r"C:\Users\brayd\Downloads\bizbhzwdqwkigpgu.csv"
     dataFile3 = r"C:\Users\brayd\Downloads\kem6kumzrbylvuz3.csv"
-    df = pd.read_csv(dataFile3)
+    df = pd.read_csv(dataFile)
 
     df['DATE'] = pd.to_datetime(df['DATE'], dayfirst=True, errors='coerce')
     df = df.sort_values(by='DATE', ascending=True)
@@ -193,20 +193,44 @@ def portfolio_returns():
     return portfolio
 
 import matplotlib.pyplot as plt
+from matplotlib.ticker import MaxNLocator
+import matplotlib.ticker as mtick
+
 def monthly_total_return():
     portfolio = portfolio_returns()
     portfolio['CUM_RET'] = portfolio.groupby(['PORTFOLIO'])['RET'].transform(lambda x: np.cumprod(1 + x))
+    portfolio = portfolio[portfolio['PORTFOLIO'].str.contains('V2')]
+    portfolio['CUM_RET'] = portfolio.groupby(['PORTFOLIO'])['CUM_RET'].transform('mean')
 
-    plt.figure(figsize=(8, 5))
-    for port in portfolio['PORTFOLIO'].unique():
-        data = portfolio[portfolio['PORTFOLIO'] == port]
-        plt.plot(data['DATE'], data['CUM_RET'], label=port)
 
-    plt.xlabel("Date")
-    plt.ylabel("Return")
-    plt.title("Monthly Portfolio Returns Over Time")
-    plt.legend()
-    plt.grid(True)
+    fig, ax = plt.subplots(figsize=(8, 5))
+    fig.patch.set_facecolor('#1D1D1B')
+    ax.set_facecolor('#1D1D1B')
+
+    ax.plot(data['DATE'], data['CUM_RET'], label='New Portfolio')
+
+    # Formatting the plot
+    ax.set_xlabel("Date", fontsize=10, color='white')
+    ax.set_ylabel("Cumulative Return", fontsize=10, color='white')
+    ax.set_title("Monthly Portfolio Returns Over Time", fontsize=12, color='white', loc='left')
+
+    ax.tick_params(axis='x', labelsize=8, colors='white')
[…]

### 2025-03-30 sumStats.py

@@ -5,6 +5,8 @@ from portfolio_monthly import portfolio_returns as monthly
 
 import pandas as pd
 import matplotlib.pyplot as plt
+from matplotlib.ticker import MaxNLocator
+import matplotlib.ticker as mtick
 import numpy as np
 
 def tickers():
@@ -21,17 +23,15 @@ def tickers():
     return data
 
 def monthly_rebalance_compare():
-
     dY = yearly()
     dM = monthly()
 
     dY = dY[['DATE', 'PORTFOLIO', 'RET']]
     dM = dM[['DATE', 'PORTFOLIO', 'RET']]
 
-
-
     d = pd.merge(dY, dM, how='left', on=['DATE', 'PORTFOLIO'])
     d = d[['DATE', 'PORTFOLIO', 'RET_x', 'RET_y']]
+    d = d[d['DATE'] >= "2007-01-01"]
     d['CUM_RET_x'] = d.groupby(['PORTFOLIO'])['RET_x'].transform(lambda x: np.cumprod(1 + x))
     d['CUM_RET_y'] = d.groupby(['PORTFOLIO'])['RET_y'].transform(lambda x: np.cumprod(1 + x))
 
@@ -40,19 +40,43 @@ def monthly_rebalance_compare():
     fig, axes = plt.subplots(2, 2, figsize=(8, 8))
     axes = axes.flatten()
 
+    fig.patch.set_facecolor('#1D1D1B')
+
     for i, portfolio in enumerate(portfolios):
         ax = axes[i]
         portfolio_data = d[d['PORTFOLIO'] == portfolio]
+
+        print(portfolio_data)
         
-        ax.plot(portfolio_data['DATE'], portfolio_data['CUM_RET_x'], label='Annual Rebalance')
-        ax.plot(portfolio_data['DATE'], portfolio_data['CUM_RET_y'], label='Monthly Rebalance')
+        ax.plot(portfolio_data['DATE'], portfolio_data['CUM_RET_x'], label='Annual Rebalance', color='#9fc3cf')
+        ax.plot(portfolio_data['DATE'], portfolio_data['CUM_RET_y'], label='Monthly Rebalance', color='#5e91a4')
+
+        ax.set_title(f'Portfolio Comparison {portfolio}', fontsize=12, color='white', loc='left')
+        ax.set_xlabel("Date", fontsize=10, color='white')
+        ax.set_ylabel("Cumulative Return", fontsize=10, color='white')
+
+        # Formatting ticks, legend, and grid
[…]

### 2025-03-30 visuals.py

@@ -6,6 +6,7 @@ import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.ticker as mtick
+from matplotlib.ticker import MaxNLocator
 from scipy.stats import norm
 
 def monthly_ret():
@@ -89,7 +90,6 @@ def monthly_total_return():
     plt.grid(True, linestyle=':', color='white')
 
     plt.show()
-
 
 
 def monthly_alpha():
@@ -227,28 +227,69 @@ def portfolio_skews():
     portfolio = pchart()
     port = portfolio[['Year', 'LongV1Skew', 'ShortV1Skew', 'LongV2Skew', 'ShortV2Skew']]
 
-    fig, axes = plt.subplots(2, 2, figsize=(8, 8), sharex=True)
+    fig, axes = plt.subplots(2, 2, figsize=(8, 8))
+    fig.patch.set_facecolor('#1D1D1B')
 
     skews = {
-        "LongV1Skew": (0, 0),
-        "ShortV1Skew": (0, 1),
-        "LongV2Skew": (1, 0),
-        "ShortV2Skew": (1, 1)
+        "LongV2Skew": (0, 0),
+        "ShortV2Skew": (0, 1)
     }
 
     for key, (row, col) in skews.items():
         ax = axes[row, col]
-        ax.plot(port["Year"], port[key], marker="o", linestyle="-", label=key)
-        ax.set_title(key)
-        ax.set_xlabel("Year ")
-        ax.set_ylabel("Skew")
-        ax.legend()
-        ax.grid(True)
+        ax.plot(port["Year"], port[key], marker="o", linestyle="-", label=key, color='#9fc3cf')
+
+        ax.set_title(key, fontsize=10, color='white', loc='left')
+        ax.set_xlabel("Year", fontsize=8, color='white')
+        ax.set_ylabel("Skew", fontsize=8, color='white')
+        ax.tick_params(axis='x', labelsize=8, colors='white')
+        ax.tick_params(axis='y', labelsize=8, colors='white')
+        ax.grid(True, linestyle=':', color='white')
+
+        ax.set_facecolor('#1D1D1B')
+        ax.spines['top'].set_color('white')
+        ax.spines['right'].set_color('white')
+        ax.spines['bottom'].set_color('white')
+        ax.spines['left'].set_color('white')
+
+        ax.xaxis.set_major_locator(MaxNLocator(integer=True))
[…]

### 2025-03-31 Animated Graph.qmd
Added Animated graph

![](images/clipboard-1227717430.png)

---
title: ""
author: "Brayden Boyko"
format: html
editor: visual
---

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(plotly)

app_df <- read.csv("C:/Users/Brayden Boyko/Downloads/322 Data CSV (1).csv")

df_long <- app_df %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "Years",
    values_to = "Value"
  ) %>%
  mutate(Years = as.numeric(str_remove(Years, "^X")))

colnames(df_long)[1] <- "Metric"

years <- sort(unique(df_long$Years))

line_df <- map_dfr(years, function(yr) {
  df_long %>%
    filter(Years <= yr) %>%
    mutate(frame = yr)
})

dot_df <- df_long %>%
  mutate(frame = Years)

p <- plot_ly()

p <- p %>%
  add_trace(
    data = line_df,
    x = ~Years,
    y = ~Value,
    frame = ~frame,
    color = ~Metric,
    type = 'scatter',
    mode = 'lines',
    line = list(width = 2),
    showlegend = TRUE
  )

p <- p %>%
  layout(
    title = "Financial Metrics Over Time",
    xaxis = list(title = "Year", color = "white"),
    yaxis = list(title = "Value (Millions)", color = "white"),
    plot_bgcolor = "#1D1D1B",
    paper_bgcolor = "#1D1D1B",
    font = list(color = "white")
  ) %>%
  animation_opts(frame = 600, easing = 'linear', redraw = FALSE) %>%
  animation_button(x = 1, xanchor = "right", y = 0, yanchor = "bottom") %>%
  animation_slider(
    currentvalue = list(prefix = "Year: ", font = list(color = 'white')),
    tickfont = list(color = 'white')
  )

p

```

### 2025-03-31 animation.py

@@ -1,0 +1,67 @@
+import pandas as pd
+import plotly.express as px
+import plotly.graph_objects as go
+
+df = pd.read_csv("C:/Users/Brayden Boyko/Downloads/322 Data CSV.csv")
+
+df = df.melt(id_vars=['Year'], value_vars=['Revenue', 'EBIT', 'NOPAT'], 
+             var_name='name', value_name='value')
+
+df = df.dropna()
+
+fig = px.line(df, 
+              x='Year', 
+              y='value', 
+              color='name', 
+              line_group='name', 
+              animation_frame='Year',  # Animation on 'Year'
+              title="Performance Metrics Over Time",
+              labels={"Year": "Years", "value": "Value", "name": "Line Item"})
+
+fig.update_layout(
+    plot_bgcolor='#222',
+    paper_bgcolor='#222',
+    font=dict(color='white'),
+    xaxis=dict(title=dict(text='Years', font=dict(color='white')), tickfont=dict(color='white')),
+    yaxis=dict(title=dict(text='', font=dict(color='white')), tickfont=dict(color='white')),
+    legend=dict(bgcolor='#222', font=dict(color='white'))
+)
+
+fig.update_traces(line=dict(width=1.2)) 
+fig.update_yaxes(tickformat="$") 
+
+fig.update_layout(
+    updatemenus=[
+        dict(
+            type='buttons',
+            showactive=False,
+            buttons=[
+                dict(label='Play',
+                     method='animate',
+                     args=[None, dict(frame=dict(duration=500, redraw=True),
+                                      fromcurrent=True)]),
+                dict(label='Pause',
+                     method='animate',
+                     args=[[None], dict(frame=dict(duration=0, redraw=False),
+                                        mode='immediate')])
+            ],
+            x=1, xanchor='right',
+            y=0, yanchor='bottom',
+            font=dict(color='white')
+        )
+    ]
+)
+
+fig.update_layout(
+    sliders=[
+        dict(
+            active=0,
+            currentvalue=dict(prefix="Year: ", font=dict(color='white')),
+            pad=dict(t=50),
+            tickfont=dict(color='white'),
+        )
+    ]
+)
+
+# Show plot
+fig.show()

### 2025-03-31 Animated Graph.qmd
Updated HTML
Updated Background

@@
-![](images/clipboard-1227717430.png)
-
 ---
 title: ""
-author: "Brayden Boyko"
-format: html
-editor: visual
+author: ""
+format: 
+  html:
+    embed-resources: true
+    page-layout: full
+    toc: false
+    grid: 
+      margin-width: 10px
+      body-width: 1000px
+    resources: 
+      - "*.js"
+      - "*.mp3"
+editor: source
 ---
+<style>
+body {
+  background-color: #1D1D1B;
+}
+</style>
 
